### 2016-11-05 Paper/Files/rsj2016.tex
[minor] add paper template files

\documentclass[a4paper]{article}

\usepackage{url}
\usepackage{cite}
\usepackage{array}
\usepackage{comment}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{rsj2016e_IS}
\usepackage[cmex10]{amsmath}
\usepackage[pdftex]{graphicx}
\usepackage[caption=false,font=footnotesize]{subfig}

\graphicspath{{./Images/}}
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}

\begin{document}
\title{Motor-skill Learning in Latent Spaces for Robotic Clothing Assistance}
\author{Nishanth Koganti$^{a,b}$, Ravi P. Joshi$^b$, Tomoya Tamei$^a$, Kazushi Ikeda$^a$, Tomohiro Shibata$^b$\\($^a$Graduate School of Information Science, Nara Institute of Science and Technology)\\($^b$Graduate School of Life Science and Systems Engineering, Kyushu Institute of Technology)}
\engtitle{}
\engauthor{}

\maketitle
\pagestyle{empty}
\thispagestyle{empty}
\setlength{\baselineskip}{4.4mm}

\section{Introduction}
\label{section:introduction}

Clothing assistance is a basic and important assistance activity in the daily life of the elderly and disabled people. However, robotic clothing assistance is still considered an open problem by most robotics researchers. Design of a robust framework involves reliable cloth state\ estimation in real-time and a motor skills learning framework that can detect and adapt to various failure scenarios. Existing studies do not explicitly handle these challenges and perform point-to-point motion planning in an offline manner for clothing assistance~\cite{yamazaki},\cite{gao}. An alternate approach is to formulate robotic clothing assistance as a reinforcement learning problem wherein the robot learns to recover from failure scenarios and adapt to new settings from experience.

Tamei \emph{et al.}~\cite{tamei} have proposed a reinforcement learning (RL) framework for clothing assistance with a dual-arm robot as the agent and a mannequin as the subject. The clothing task was to cloth the mannequin with a T-shirt which is initially on the mannequin's hands. The framework was formulated using low dimensional representations for the state and policy. The state was represented using topology coordinates which captured the relationship between the human and cloth extremities. The dual-arm robot's trajectory was parametrized using via-points which were obtained using a minimum jerk criterion. The policy was initialized using kinesthetic demonstration for a particular posture and the framework could successfully learn a policy for an unseen posture.

A limitation in the above framework is that policy search is performed in the kinematic space which is considerably high dimensional for a 7 degree of freedom (DOF) dual-arm robot. For tractable learning time, policy update was done using finite difference policy gradient applied to a single via-point of a single joint in each robot arm. This severely constrained the generalization capability to very different environmental settings such as major changes in the subject's posture or using different clothing article. To address this problem, we consider an alternate representation that is more flexible and suitable for robust learning of motor-skills. We propose the use of dimensionality reduction (DR) to learn a low-dimensional latent space that encodes clothing skills without losing any generalization capability.

In this study, we propose an efficient representation of motor skills that relies on the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM)~\cite{bgplvm}. BGPLVM is capable of learning a data-efficient latent space for clothing tasks performed by a dual-arm robot. Representation of clothing skills in a low-dimensional space enables the use of expressive policy update rules for generalization to very different settings. We apply our proposed method to a clothing assistance setup as shown in Figure~\ref{figure:setup}. We demonstrate that the learned space generates robot trajectories that maintain task space constraints required for clothing tasks. We further present the design of a real-time controller from a BGPLVM latent space to a dual-arm robot. The experimental results indicate a promising policy representation with reinforcement learning that can be used for robotic clothing assistance.

The rest of the paper is structured as follows. Section~\ref{section:literature} provides an overview of related studies. The proposed framework is presented in Section~\ref{section:method}. Section~\ref{section:results} includes experimental results and Section~\ref{section:conclusion} concludes the paper with directions for future work.

\section{Related Work}
\label{section:literature}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.45\textwidth, height=0.2\textheight]{setup.png}
  \caption{Clothing Assistance framework: Dual-arm robot clothing soft-mannequin with T-shirt.}
  \label{figure:setup}
\end{figure}

Motor skills learning for robots involves an agent interacting with a partially observable environment in continuous state and action spaces. Usually policy search reinforcement learning is applied wherein an optimal policy is learned by searching within the policy space. Reinforcement Learning usually suffer from the curse of dimensionality when applied to real-world situations as the search space grows exponentially. Existing frameworks do not scale well to high-dimensionality and usually take a long learning time without human-specified constraints. An efficient approach is the use of dimensionality reduction (DR) along with reinforcement learning (RL). The combination of DR leads to a tractable search space for RL resulting in a faster learning time. There have been several studies in this direction in the recent years.

Bitzer \emph{et al.}~\cite{bitzer} proposed a RL framework wherein DR was used as a preprocessing step to extract the problem structure from task-specific robot postures. Gaussian Process Latent Variable Model (GPLVM), was used to capture the task space constraints in a low-dimensional latent representation and Temporal Difference learning (TD(0)) was applied in the learned space. This framework was applied to bi-manual reaching task on a full-humanoid robot (19 DOF). Experimental results demonstrated that combining DR and RL significantly outperforms using RL in the full state space. However, GPLVM relies on a maximum-a-posteriori (MAP) estimate of the latent space and can overfit to the training data.

Luck \emph{et al.}~\cite{luck} proposed a policy search framework for robotics that inherently combines RL and DR wherein the latent space learning is also based on the reward signals during learning. The framework relies on an Expectation-Maximization (EM) framework where in the parameters for the policy representation as well as the projection parameters for the latent space are learned at the same time. This framework was applied for standing on a single leg task for a full-humanoid robot and was able to quickly learn an optimal policy without requiring any initial demonstrations. However, use of linear dimensionality reduction constrains the strength of the latent space search especially for non-linear tasks such as solving inverse kinematics.

In this study, we propose the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM) to learn a low-dimensional latent space for encoding clothing skills. The advantage of BGPLVM is that it relies on variational inference to learn a posterior distribution on the latent space rather than a MAP estimate as in GPLVM. This avoids over fitting to the training data, thereby, improving the generalization capability of the model to unseen environmental settings. We further explore various representations to the learning of BGPLVM model specific to clothing assistance tasks.

\section{Proposed Method}
\label{section:method}

In this study, we explore the suitability of using Bayesian Gaussian Process Latent Variable Model (BGPLVM) to learn a latent space as a low dimensional representation of motor skills for clothing assistance. This section is divided into two parts. Section~\ref{section:bgplvm} provides the formulation of BGPLVM and Section~\ref{section:clothassist} presents the application of BGPLVM in the clothing assistance framework.

\subsection{Bayesian Latent Space Learning}
\label{section:bgplvm}

Bayesian Gaussian Process Latent Variable Model (BGPLVM) is a dimensionality reduction technique proposed by Titsias \emph{et al.}~\cite{bgplvm}. BGPLVM is derived from the probabilistic model where the observations, $\mathbf{Y} = \{\mathbf{y}_{i} \in \mathbb{R}^D\}_{n=1}^N$, are assumed to be generated from latent inputs $\mathbf{X} = \{\mathbf{x}_{i} \in \mathbb{R}^q\}_{n=1}^N$ through a noisy process,
\begin{equation}
  \mathbf{y}_i = f(\mathbf{x}_i) + \epsilon,~\epsilon \sim \mathcal{N}(\mathbf{0},\beta^{-1}\mathbf{I})
\end{equation}
The mapping function $f$ is modeled using a Gaussian Process (GP) which makes the model capable of performing non-linear dimensionality reduction with the use of a non-linear kernel function. The marginal likelihood for the generative model is given by:
\begin{equation}
  \label{eqn:marginallikelihood}
  p(\mathbf{Y}|\mathbf{X},\Phi) = \prod_{d=1}^{D} \mathcal{N}(\mathbf{y}_{:,d}|\mathbf{0},\mathbf{K}+\beta^{-1}I)\\
\end{equation}
where $\mathbf{X},\Phi$ are the unknown latent positions and hyper parameters for the GP mapping that need to be inferred. $\mathbf{K}$ is the kernel matrix constructed from the latent points.

In the generative model, the latent positions need to be marginalized out for having a purely Bayesian treatment:
\begin{equation}
  \label{eqn:marginalization}
  p(\mathbf{Y}|\Phi) = \int p(\mathbf{Y}|\mathbf{X},\Phi) p(\mathbf{X}) d\mathbf{X}
\end{equation}
However, the integral becomes intractable as $\mathbf{X}$ appears non-linearly in the kernel covariance matrix for a non-linear kernel function as shown in Equation~(\ref{eqn:marginallikelihood}). Titsias \emph{et al.}~\cite{bgplvm} proposed a variational inference approach to compute a tractable lower bound for the marginalization thereby inferring a posterior distribution on the latent positions rather than a MAP estimate. Detailed derivations of the model are further presented in \cite{bgplvm}. For performing automatic model selection of the latent space dimensionality, the Automatic Relevance Determination (ARD) Kernel can be used in the GP mapping:
\begin{equation}
  \label{eqn:ardkernel}
  k_{\text{ard}}(\mathbf{x}_i,\mathbf{x}_j) = \sigma_{ard}^2~\text{exp}\left( - \frac{1}{2} \sum_{k=1}^q{\alpha_q (x_{i,k} - x_{j,k})^2}\right)
\end{equation}
The ARD weights $\alpha_q$ describe the relevance of each dimension and zero weight indicating complete irrelevance. Maximizing the marginal likelihood w.r.t. these weights allows the inference of latent space dimensionality.

The inference for unseen test data can now be performed through a Bayesian formulation instead of relying on a MAP estimation of the latent space. The predictive distribution is given by the ratio of two marginal likelihoods, both of which can be approximated using the variational inference technique:
\begin{equation}
	\label{eqn:testinference}
	p(\mathbf{y}^*|\mathbf{Y}) = \frac{\int p(\mathbf{y}^*,\mathbf{Y}|\mathbf{x}^*,\mathbf{X})p(\mathbf{x}^*,\mathbf{X})d\mathbf{X}d\mathbf{x}^*}{\int p(\mathbf{Y}|\mathbf{X})p(\mathbf{X})d\mathbf{X}}
\end{equation}
Efficient computations to handle test data is further described in \cite{bgplvm}.

\subsection{Representation of Clothing Skills using BGPLVM}
\label{section:clothassist}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{controller.png}
  \caption{Overview of latent space controller for performing clothing tasks.}
  \label{figure:realtime}
\end{figure}

Motor skills for clothing assistance lie in an high dimensional kinematic space given by joint angle trajectories of a 7 DOF dual-arm robot. The robot also has to maintain several task space constraints such as the coupling with a clothing article along with operating safely in close proximity to a human subject as shown in Figure~\ref{figure:setup}. These task space constraints are difficult to be modeled in the kinematic space as they are non-linearly related to each other. To address these problems, we propose the use of BGPLVM for learning a latent space that efficiently encodes clothing assistance skills. BGPLVM results in the learning of a low-dimensional latent space through a nonlinear and data-efficient mapping to the kinematic space. Furthermore, we train the model using successful trajectories of the robot performing clothing tasks so that the resultant latent space generates robot trajectories that follow task space constraints. The Bayesian treatment avoids over fitting to the training data and automatic inference of the latent space dimensionality through the use of ARD kernel.

In this study, we consider the clothing task where a dual-arm robot clothes a soft mannequin with a T-shirt which is initially resting on the mannequin's arms. The purpose of the BGPLVM model is to learn a latent representation $\mathbf{X} = [\mathbf{X}_1, \cdots, \mathbf{X}_N]$ from a dataset of 7-DOF dual-arm robot poses $\mathbf{Y} = [\mathbf{y}_1, \cdots, \mathbf{y}_N]$. We represent the robot poses in two alternate ways i.e. 1) Kinematic representation given by joint angles of both 7-DOF arms $D_K = 14$ and 2) Task space representation given by the end-effector pose of both robot arms which comprises the Cartesian position ${P_X,P_Y,P_Z} \in \mathbb{R}^3$ and orientation represented using a quaternion ${O_X,O_Y,O_Z,O_{\omega}} \in \mathbb{R}^4$ forming a 14-dimensional space $D_T = 14$. The dataset $\mathbf{Y}$ was obtained from a collection of successful clothing assistance demonstrations performed by the robot for various postures of the mannequin. We set the dimension of the latent space as $q = 5$, however the dimensionality is eventually inferred through the training of the ARD kernel weights as explained in Section~\ref{section:bgplvm}.

\subsection{Latent Space Robot Controller}
\label{section:controller}

There can be several types of failure scenarios when the robot performs clothing tasks. While clothing, the T-shirt collar could get stuck on the mannequin's head and while unclothing, the T-shirt could get stuck in the mannequin's shoulder joints. To recover from these failures, not only is the trajectory of the robot important, but also the speed of execution. Imparting these skills through kinesthetic movement of the arms can be difficult for a bulky robot and could lead to noisy demonstrations. To address this problem, we have implemented a real-time controller that gets an input signal from the BGPLVM latent space. The BGPLVM model learns a mapping from the low dimensional latent space to the robot kinematic space such that a trajectory of latent points generates a trajectory on the dual-arm robot. This interface can be used as a tool for Learning from Demonstration (LfD) wherein, the necessary clothing skills are imparted to the robot by using cursor control over the latent space as shown in Figure~\ref{figure:realtime}.

\section{Results}
\label{section:results}

In this section, we present the results of using BGPLVM as a representation for clothing assistance motor skills. The experimental setup includes the Baxter 7 DOF dual-arm robot performing clothing tasks and a soft mannequin as the subject as shown in Figure~\ref{figure:setup}. The clothing task is to cloth the mannequin with a T-shirt which is initially resting on the mannequin's arms. The evaluation is presented in two parts: Section~\ref{section:performance} demonstrates the generalization capability of BGPLVM to unseen environmental settings and Section~\ref{section:controllerres} presents the performance of the real-time controller presented in Section~\ref{section:controller}.

\subsection{Predictive Performance}
\label{section:performance}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{latentspace.png}
  \caption{Latent Space of BGPLVM trained on Joint Angle dataset of clothing tasks. Latent points include 3 train trials and 1 test trial.}
  \label{figure:latent}
\end{figure}

In this experiment, the generalization capability of BGPLVM for performing clothing tasks is evaluated. The evaluation dataset contains 4 clothing trials that were obtained for 4 different postures of the mannequin wherein the elevation of the arms were varied forming different angles $\{65^o, 70^o, 75^o, 80^o\}$ with the mannequin's body. The clothing trial for $75^o$ was left out as the test data and BGPLVM models were trained for the remaining 3 clothing trials. Trained models for both the joint angles (JA) and end-effector Cartesian positions (EE) resulted in 2 dimensional latent spaces. The resultant latent space for the JA model is visualized in Figure~\ref{figure:latent}. It can be seen that smooth latent trajectories are formed for each clothing trial capturing the dynamics of performing the task. Furthermore, a smooth latent trajectory exists for the test clothing trial as well indicating the possibility for a reinforcement learning framework to generalize to the new posture.

\subsection{Controller Demonstration}
\label{section:controllerres}

In this section, the latent space controller presented in Section~\ref{section:controller} is demonstrated. The latent space for the controller is trained for the joint angles of a clothing trial obtained from kinesthetic demonstration by a human. The resultant model represented the data in a 2 dimensional latent space. Cursor control on the latent space was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points as shown in Figure~\ref{figure:controller}. The resultant latent dimensions further captured a specific aspect of the clothing motor-skills. For example, the most significant dimension captured the horizontal motion of the arms along the mannequin while maintaining the constraints for clothing. The second dimension captured various vertical motions of pulling up the T-shirt in the beginning and pulling it down along the torso at the end.

\section{Conclusion}
\label{section:conclusion}

Motor-skills learning for robotic clothing assistance task involves a high dimensional kinematic search space and maintaining several task space constraints. In this study, we have presented the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM) as a representation for encoding motor-skills to perform clothing assistance task. The experimental results indicate our method as a promising approach in combination with reinforcement learning. A latent space representation can be learned from multiple observation spaces using Manifold Relevance Determination (MRD)~\cite{mrd} which is an extension of BGPLVM. Based on this flexibility, our future work will be to learn models that explicitly incorporates visual information of the relationship between the human and cloth. The long term goal will be to develop a policy search reinforcement learning framework that relies on the use of BGPLVM.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{controller2.png}
  \caption{Cursor control on BGPLVM latent space to perform clothing task. Figure on left shows simulation of robot pose, right top shows robot performing clothing task and right bottom shows cursor control in latent space.}
  \label{figure:controller}
\end{figure}

\section*{Acknowledgement}
This work was supported in part by ImPACT Program of Council for Science, Technology and Innovation (Cabinet Office, Government of Japan) 2015-PM07-36-01. This work was also supported in part by the Grant-in-Aid for Scientific Research from Japan Society for the Promotion of Science (No. 16H01749).

\small{
\begin{thebibliography}{15}
\bibitem{yamazaki}
K. Yamazaki, R. Oya, K. Nagahama, K. Okada, and M. Inaba, ``Bottom dressing by a life-sized humanoid robot provided failure detection and recovery functions'' in Proc. of the IEEE Intl. Conf. on System Integration (SII), 2014.

\bibitem{gao}
Y. Gao, H. J. Chang, and Y. Demiris, ``User modelling for personalised dressing assistance by humanoid robots'' in Proc of the IEEE Intl. Conf. on Intelligent Robots and Systems (IROS), 2015.

\bibitem{tamei}
T. Tamei, T. Matsubara, A. Rai, and T. Shibata, ``Reinforcement Learning of Clothing Assistance with a Dual-arm Robot'' in Proc. of the IEEE-RAS Intl. Conf. on Humanoid Robots (Humanoids), 2011.

\bibitem{bgplvm}
Titsias, M. K., and Lawrence, N. D, ``Bayesian Gaussian process latent variable model'' in Proc. of the Intl. Conf. on Artificial Intelligence and Statistics, 2010.

\bibitem{bitzer}
Bitzer, S., Howard, M., and Vijayakumar, S, ``Using dimensionality reduction to exploit constraints in reinforcement learning'' in Proc. of the IEEE Intl. Conf. on Intelligent Robots and Systems (IROS), 2010.

\bibitem{luck}
Luck, K. S., Neumann, G., Berger, E., Peters, J., and Ben Amor, H, ``Latent space policy search for robotics'' in Proc. of the IEEE Intl. Conf. on Intelligent Robots and Systems (IROS), 2014.

\bibitem{mrd}
Damianou, A., Ek, C., Titsias, M., and Lawrence, N, ``Manifold relevance determination'' in Proc. of the Intl. Conf. on Machine Learning, 2012.

\end{thebibliography}
}
\normalsize

\end{document}

### 2016-11-05 Paper/paper.tex
[major] first draft of paper based on rsj paper

--- 2016-11-05 Paper/Files/rsj2016.tex
@@
-\documentclass[a4paper]{article}
+\documentclass{article}
+
+\usepackage[final,nonatbib]{nips_2016}
 
 \usepackage{url}
-\usepackage{cite}
-\usepackage{array}
-\usepackage{comment}
-\usepackage{multirow}
+\usepackage{amsmath}
 \usepackage{booktabs}
+\usepackage{hyperref}
 \usepackage{amsfonts}
-\usepackage{rsj2016e_IS}
-\usepackage[cmex10]{amsmath}
+\usepackage{nicefrac}
+\usepackage{microtype}
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
 \usepackage[pdftex]{graphicx}
@@
 
+\title{Data-efficient Motor-skills Learning in Latent Spaces for Robotic Clothing Assistance}
+
+\author{
+  Nishanth Koganti$^{1,2}$, Tomoya Tamei$^{1}$, Kazushi Ikeda$^{1}$, Tomohiro Shibata$^{2}$\\
+  $^1$ Nara Institute of Science and Technology, Japan~~~$^2$ Kyushu Institute of Technology, Japan\\
+  \texttt{\{nishanth-k,tomo-tam,kazushi\}@is.naist.jp, tom@brain.kyutech.ac.jp} \\
+}
+
 \begin{document}
-\title{Motor-skill Learning in Latent Spaces for Robotic Clothing Assistance}
-\author{Nishanth Koganti$^{a,b}$, Ravi P.
-Joshi$^b$, Tomoya Tamei$^a$, Kazushi Ikeda$^a$, Tomohiro Shibata$^b$\\($^a$Graduate School of Information Science, Nara Institute of Science and Technology)\\($^b$Graduate School of Life Science and Systems Engineering, Kyushu Institute of Technology)}
-\engtitle{}
-\engauthor{}
 
 \maketitle
-\pagestyle{empty}
-\thispagestyle{empty}
-\setlength{\baselineskip}{4.4mm}
+
+\begin{abstract}
+Robotic implementation of clothing assistance can greatly improve the quality-of-life for the elderly and disabled.
+This is still considered an open problem as it involves close interaction of the robot with non-rigid clothing articles and the assisted person with varying posture.
+In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent manifold learning.
+We implement our framework in a practical setting with a dual-arm robot performing clothing tasks.
+We demonstrate that performing motor-skills learning in such task specific latent spaces outperforms learning in the high-dimensional joint configuration space of the robot.
+Furthermore, our framework can also be used as a tool for learning from demonstration to impart novel skills to the robot.
+\end{abstract}
 
@@
 
-\small{
+\small
+
 \begin{thebibliography}{15}
@@
 \end{thebibliography}
-}
-\normalsize
 

### 2016-11-09 Paper/paper.tex
[major] reduce content of paper

@@
 \begin{abstract}
-Robotic implementation of clothing assistance can greatly improve the quality-of-life for the elderly and disabled.
-This is still considered an open problem as it involves close interaction of the robot with non-rigid clothing articles and the assisted person with varying posture.
+Robotic implementation of clothing assistance is a challenging problem that can greatly improve the quality-of-life for the elderly and disabled.
 In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent manifold learning.
@@
 
-Clothing assistance is a basic and important assistance activity in the daily life of the elderly and disabled people.
-However, robotic clothing assistance is still considered an open problem by most robotics researchers.
-Design of a robust framework involves reliable cloth state\ estimation in real-time and a motor skills learning framework that can detect and adapt to various failure scenarios.
-Existing studies do not explicitly handle these challenges and perform point-to-point motion planning in an offline manner for clothing assistance~\cite{yamazaki},\cite{gao}.
-An alternate approach is to formulate robotic clothing assistance as a reinforcement learning problem wherein the robot learns to recover from failure scenarios and adapt to new settings from experience.
-
-Tamei \emph{et al.}~\cite{tamei} have proposed a reinforcement learning (RL) framework for clothing assistance with a dual-arm robot as the agent and a mannequin as the subject.
-The clothing task was to cloth the mannequin with a T-shirt which is initially on the mannequin's hands.
-The framework was formulated using low dimensional representations for the state and policy.
-The state was represented using topology coordinates which captured the relationship between the human and cloth extremities.
-The dual-arm robot's trajectory was parametrized using via-points which were obtained using a minimum jerk criterion.
-The policy was initialized using kinesthetic demonstration for a particular posture and the framework could successfully learn a policy for an unseen posture.
-
-A limitation in the above framework is that policy search is performed in the kinematic space which is considerably high dimensional for a 7 degree of freedom (DOF) dual-arm robot.
+Clothing assistance is a necessity in the daily life of the elderly and disabled people.
+However, robotic clothing assistance is still considered an open problem.
+Design of a robust framework involves reliable cloth state estimation in real-time and a motor skills learning framework that can detect and adapt to various failure scenarios.
+Tamei \emph{et al.}~\cite{tamei} proposed a reinforcement learning (RL) framework for clothing assistance with a dual-arm robot to cloth a soft mannequin with a T-shirt initially resting on its arms.
 For tractable learning time, policy update was done using finite difference policy gradient applied to a single via-point of a single joint in each robot arm.
 This severely constrained the generalization capability to very different environmental settings such as major changes in the subject's posture or using different clothing article.
-To address this problem, we consider an alternate representation that is more flexible and suitable for robust learning of motor-skills.
-We propose the use of dimensionality reduction (DR) to learn a low-dimensional latent space that encodes clothing skills without losing any generalization capability.
+
+To address this problem, we need to consider an alternate representation that is more flexible and suitable for robust learning of motor-skills.
+An efficient approach is the use of dimensionality reduction (DR) along with reinforcement learning (RL).
+There have been several studies in this direction in the recent years.
+Bitzer \emph{et al.}~\cite{bitzer} proposed a learning framework where DR was used as a preprocessing step to extract the problem structure.
+However, this study relies on a maximum-a-posteriori (MAP) estimate of the latent space and can overfit to the training data.
+Luck \emph{et al.}~\cite{luck} proposed a policy search framework that inherently combines RL and DR wherein the latent space is learned based on the reward signals.
+However, this framework relies on the use of linear dimensionality reduction which limits applicability to complex tasks.
 
@@
 We apply our proposed method to a clothing assistance setup as shown in Figure~\ref{figure:setup}.
-We demonstrate that the learned space generates robot trajectories that maintain task space constraints required for clothing tasks.
-We further present the design of a real-time controller from a BGPLVM latent space to a dual-arm robot.
-The experimental results indicate a promising policy representation with reinforcement learning that can be used for robotic clothing assistance.
+We demonstrate that performing policy search in this latent space outperforms learning in the high-dimensional action space.
+We further present the design of a real-time controller from the BGPLVM latent space that can be used as a tool for Learning from Demonstration (LfD).
+The experimental results indicate a promising reinforcement learning learning that can be used for challenging tasks such as robotic clothing assistance.
 
 The rest of the paper is structured as follows.
-Section~\ref{section:literature} provides an overview of related studies.
 The proposed framework is presented in Section~\ref{section:method}.
 Section~\ref{section:results} includes experimental results and Section~\ref{section:conclusion} concludes the paper with directions for future work.
-
-\section{Related Work}
-\label{section:literature}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.45\textwidth, height=0.2\textheight]{setup.png}
-  \caption{Clothing Assistance framework: Dual-arm robot clothing soft-mannequin with T-shirt.}
-  \label{figure:setup}
-\end{figure}
-
-Motor skills learning for robots involves an agent interacting with a partially observable environment in continuous state and action spaces.
-Usually policy search reinforcement learning is applied wherein an optimal policy is learned by searching within the policy space.
-Reinforcement Learning usually suffer from the curse of dimensionality when applied to real-world situations as the search space grows exponentially.
-Existing frameworks do not scale well to high-dimensionality and usually take a long learning time without human-specified constraints.
-An efficient approach is the use of dimensionality reduction (DR) along with reinforcement learning (RL).
-The combination of DR leads to a tractable search space for RL resulting in a faster learning time.
-There have been several studies in this direction in the recent years.
-
-Bitzer \emph{et al.}~\cite{bitzer} proposed a RL framework wherein DR was used as a preprocessing step to extract the problem structure from task-specific robot postures.
-Gaussian Process Latent Variable Model (GPLVM), was used to capture the task space constraints in a low-dimensional latent representation and Temporal Difference learning (TD(0)) was applied in the learned space.
-This framework was applied to bi-manual reaching task on a full-humanoid robot (19 DOF).
-Experimental results demonstrated that combining DR and RL significantly outperforms using RL in the full state space.
-However, GPLVM relies on a maximum-a-posteriori (MAP) estimate of the latent space and can overfit to the training data.
-
-Luck \emph{et al.}~\cite{luck} proposed a policy search framework for robotics that inherently combines RL and DR wherein the latent space learning is also based on the reward signals during learning.
-The framework relies on an Expectation-Maximization (EM) framework where in the parameters for the policy representation as well as the projection parameters for the latent space are learned at the same time.
-This framework was applied for standing on a single leg task for a full-humanoid robot and was able to quickly learn an optimal policy without requiring any initial demonstrations.
-However, use of linear dimensionality reduction constrains the strength of the latent space search especially for non-linear tasks such as solving inverse kinematics.
-
-In this study, we propose the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM) to learn a low-dimensional latent space for encoding clothing skills.
-The advantage of BGPLVM is that it relies on variational inference to learn a posterior distribution on the latent space rather than a MAP estimate as in GPLVM.
-This avoids over fitting to the training data, thereby, improving the generalization capability of the model to unseen environmental settings.
-We further explore various representations to the learning of BGPLVM model specific to clothing assistance tasks.
 
@@
 This section is divided into two parts.
-Section~\ref{section:bgplvm} provides the formulation of BGPLVM and Section~\ref{section:clothassist} presents the application of BGPLVM in the clothing assistance framework.
+Section~\ref{section:bgplvm} provides the formulation of BGPLVM and Sections~\ref{section:clothassist}, ref{section:controller} presents the application of BGPLVM in the clothing assistance framework.
 
-\subsection{Bayesian Latent Space Learning}
+\subsection{Bayesian Nonparametric Latent Space Learning}
 \label{section:bgplvm}
 
-Bayesian Gaussian Process Latent Variable Model (BGPLVM) is a dimensionality reduction technique proposed by Titsias \emph{et al.}~\cite{bgplvm}.
-BGPLVM is derived from the probabilistic model where the observations, $\mathbf{Y} = \{\mathbf{y}_{i} \in \mathbb{R}^D\}_{n=1}^N$, are assumed to be generated from latent inputs $\mathbf{X} = \{\mathbf{x}_{i} \in \mathbb{R}^q\}_{n=1}^N$ through a noisy process,
+Bayesian Gaussian Process Latent Variable Model (BGPLVM) is a dimensionality reduction technique proposed by Titsias \emph{et al.}~\cite{bgplvm}  derived from the probabilistic model where the observations, $\mathbf{Y} = \{\mathbf{y}_{i} \in \mathbb{R}^D\}_{n=1}^N$, are assumed to be generated from latent inputs $\mathbf{X} = \{\mathbf{x}_{i} \in \mathbb{R}^q\}_{n=1}^N$ through a noisy process,
 \begin{equation}
@@
 \end{equation}
-The mapping function $f$ is modeled using a Gaussian Process (GP) which makes the model capable of performing non-linear dimensionality reduction with the use of a non-linear kernel function.
+The mapping function $f$ is modeled using a Gaussian Process (GP) suitable for performing non-linear dimensionality reduction with use of a non-linear kernel function.
 The marginal likelihood for the generative model is given by:
@@
 \end{equation}
-where $\mathbf{X},\Phi$ are the unknown latent positions and hyper parameters for the GP mapping that need to be inferred. $\mathbf{K}$ is the kernel matrix constructed from the latent points.
+where $\mathbf{X},\Phi$ are the unknown latent points and hyper parameters for the GP mapping that need to be inferred.
+$\mathbf{K}$ is the kernel matrix constructed from the latent points.
+In the generative model, the latent positions need to be marginalized out for having a purely Bayesian treatment.
+Titsias \emph{et al.}~\cite{bgplvm} proposed a variational inference approach to compute a tractable lower bound for the marginalization thereby inferring a posterior distribution on the latent positions rather than a MAP estimate.
 
-In the generative model, the latent positions need to be marginalized out for having a purely Bayesian treatment:
-\begin{equation}
-  \label{eqn:marginalization}
-  p(\mathbf{Y}|\Phi) = \int p(\mathbf{Y}|\mathbf{X},\Phi) p(\mathbf{X}) d\mathbf{X}
-\end{equation}
-However, the integral becomes intractable as $\mathbf{X}$ appears non-linearly in the kernel covariance matrix for a non-linear kernel function as shown in Equation~(\ref{eqn:marginallikelihood}).
-Titsias \emph{et al.}~\cite{bgplvm} proposed a variational inference approach to compute a tractable lower bound for the marginalization thereby inferring a posterior distribution on the latent positions rather than a MAP estimate.
-Detailed derivations of the model are further presented in \cite{bgplvm}.
 For performing automatic model selection of the latent space dimensionality, the Automatic Relevance Determination (ARD) Kernel can be used in the GP mapping:
@@
 \end{equation}
-The ARD weights $\alpha_q$ describe the relevance of each dimension and zero weight indicating complete irrelevance.
+The ARD weights $\alpha_q$ describe the relevance of each dimension and zero weight indicates complete irrelevance.
 Maximizing the marginal likelihood w.r.t.
 these weights allows the inference of latent space dimensionality.
-
 The inference for unseen test data can now be performed through a Bayesian formulation instead of relying on a MAP estimation of the latent space.
@@
 
-\subsection{Representation of Clothing Skills using BGPLVM}
-\label{section:clothassist}
+\subsection{Motor Skills Representation using BGPLVM}
+\label{section:controller}
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.5\textwidth]{controller.png}
-  \caption{Overview of latent space controller for performing clothing tasks.}
-  \label{figure:realtime}
-\end{figure}
-
-Motor skills for clothing assistance lie in an high dimensional kinematic space given by joint angle trajectories of a 7 DOF dual-arm robot.
-The robot also has to maintain several task space constraints such as the coupling with a clothing article along with operating safely in close proximity to a human subject as shown in Figure~\ref{figure:setup}.
-These task space constraints are difficult to be modeled in the kinematic space as they are non-linearly related to each other.
-To address these problems, we propose the use of BGPLVM for learning a latent space that efficiently encodes clothing assistance skills.
-BGPLVM results in the learning of a low-dimensional latent space through a nonlinear and data-efficient mapping to the kinematic space.
-Furthermore, we train the model using successful trajectories of the robot performing clothing tasks so that the resultant latent space generates robot trajectories that follow task space constraints.
-The Bayesian treatment avoids over fitting to the training data and automatic inference of the latent space dimensionality through the use of ARD kernel.
-
-In this study, we consider the clothing task where a dual-arm robot clothes a soft mannequin with a T-shirt which is initially resting on the mannequin's arms.
-The purpose of the BGPLVM model is to learn a latent representation $\mathbf{X} = [\mathbf{X}_1, \cdots, \mathbf{X}_N]$ from a dataset of 7-DOF dual-arm robot poses $\mathbf{Y} = [\mathbf{y}_1, \cdots, \mathbf{y}_N]$.
-We represent the robot poses in two alternate ways i.e.
-1) Kinematic representation given by joint angles of both 7-DOF arms $D_K = 14$ and 2) Task space representation given by the end-effector pose of both robot arms which comprises the Cartesian position ${P_X,P_Y,P_Z} \in \mathbb{R}^3$ and orientation represented using a quaternion ${O_X,O_Y,O_Z,O_{\omega}} \in \mathbb{R}^4$ forming a 14-dimensional space $D_T = 14$.
-The dataset $\mathbf{Y}$ was obtained from a collection of successful clothing assistance demonstrations performed by the robot for various postures of the mannequin.
+Motor skills for clothing assistance are given by high dimensional joint angle trajectories of a 14 DOF dual-arm robot.
+The robot also has to maintain several task space constraints such as coupling with a clothing article along with safe human-robot interaction as shown in Figure~\ref{figure:setup}.
+To address these problems, we propose the use of BGPLVM for learning a low-dimensional latent space through a nonlinear mapping to the kinematic space.
+The Bayesian treatment avoids over fitting to the training data and automatic inference of latent space dimensionality through the ARD kernel.
+We consider the clothing task where a dual-arm robot clothes a soft mannequin with a T-shirt which is initially resting on the mannequin's arms.
+We represent the robot poses in two ways i.e.
+1) Kinematic representation given by joint angles $D_K = 14$ and 2) Task space representation given by the end-effector pose of both arms with cartesian position ${P_X,P_Y,P_Z} \in \mathbb{R}^3$ and orientation ${O_X,O_Y,O_Z,O_{\omega}} \in \mathbb{R}^4$ forming a 14-dimensional space $D_T = 14$.
 We set the dimension of the latent space as $q = 5$, however the dimensionality is eventually inferred through the training of the ARD kernel weights as explained in Section~\ref{section:bgplvm}.
 
-\subsection{Latent Space Robot Controller}
-\label{section:controller}
-
+The dataset, $\mathbf{Y}$, was obtained from successful clothing assistance demonstrations for various postures of the mannequin.
 There can be several types of failure scenarios when the robot performs clothing tasks.
@@
 The BGPLVM model learns a mapping from the low dimensional latent space to the robot kinematic space such that a trajectory of latent points generates a trajectory on the dual-arm robot.
-This interface can be used as a tool for Learning from Demonstration (LfD) wherein, the necessary clothing skills are imparted to the robot by using cursor control over the latent space as shown in Figure~\ref{figure:realtime}.
+This interface can be used as a tool for Learning from Demonstration (LfD) where the necessary clothing skills are imparted to the robot by using cursor control over the latent space as shown in Figure~\ref{figure:realtime}.
+
+\subsection{Latent Space Reinforcement Learning}
+\label{section:latentspace}
+
+The latent space learned using BGPLVM is an efficient search space to learn policies that generalize well to unseen environmental settings.
 
@@
 In this section, we present the results of using BGPLVM as a representation for clothing assistance motor skills.
-The experimental setup includes the Baxter 7 DOF dual-arm robot performing clothing tasks and a soft mannequin as the subject as shown in Figure~\ref{figure:setup}.
+The experimental setup includes the Baxter 14 DOF dual-arm robot performing clothing tasks and a soft mannequin as the subject as shown in Figure~\ref{figure:setup}.
 The clothing task is to cloth the mannequin with a T-shirt which is initially resting on the mannequin's arms.
 The evaluation is presented in two parts: Section~\ref{section:performance} demonstrates the generalization capability of BGPLVM to unseen environmental settings and Section~\ref{section:controllerres} presents the performance of the real-time controller presented in Section~\ref{section:controller}.
-
-\subsection{Predictive Performance}
-\label{section:performance}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.5\textwidth]{latentspace.png}
-  \caption{Latent Space of BGPLVM trained on Joint Angle dataset of clothing tasks. Latent points include 3 train trials and 1 test trial.}
-  \label{figure:latent}
-\end{figure}
-
-In this experiment, the generalization capability of BGPLVM for performing clothing tasks is evaluated.
-The evaluation dataset contains 4 clothing trials that were obtained for 4 different postures of the mannequin wherein the elevation of the arms were varied forming different angles $\{65^o, 70^o, 75^o, 80^o\}$ with the mannequin's body.
-The clothing trial for $75^o$ was left out as the test data and BGPLVM models were trained for the remaining 3 clothing trials.
-Trained models for both the joint angles (JA) and end-effector Cartesian positions (EE) resulted in 2 dimensional latent spaces.
-The resultant latent space for the JA model is visualized in Figure~\ref{figure:latent}.
-It can be seen that smooth latent trajectories are formed for each clothing trial capturing the dynamics of performing the task.
-Furthermore, a smooth latent trajectory exists for the test clothing trial as well indicating the possibility for a reinforcement learning framework to generalize to the new posture.
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.5\textwidth]{controller2.png}
-  \caption{Cursor control on BGPLVM latent space to perform clothing task.
-Figure on left shows simulation of robot pose, right top shows robot performing clothing task and right bottom shows cursor control in latent space.}
-  \label{figure:controller}
-\end{figure}
-
 \section*{Acknowledgement}
@@
 \begin{thebibliography}{15}
-\bibitem{yamazaki}
-K.
-Yamazaki, R.
-Oya, K.
-Nagahama, K.
-Okada, and M.
-Inaba, ``Bottom dressing by a life-sized humanoid robot provided failure detection and recovery functions'' in Proc.
-of the IEEE Intl.
-Conf.
-on System Integration (SII), 2014.
-
-\bibitem{gao}
-Y. Gao, H. J. Chang, and Y. Demiris, ``User modelling for personalised dressing assistance by humanoid robots'' in Proc of the IEEE Intl. Conf. on Intelligent Robots and Systems (IROS), 2015.
-
 \bibitem{tamei}

### 2016-11-09 Paper/paper.tex
[major] complete first draft of paper

@@
 Representation of clothing skills in a low-dimensional space enables the use of expressive policy update rules for generalization to very different settings.
-We apply our proposed method to a clothing assistance setup as shown in Figure~\ref{figure:setup}.
+We apply our proposed method to a clothing assistance setup as shown in Figure~\ref{figure:overview}.
 We demonstrate that performing policy search in this latent space outperforms learning in the high-dimensional action space.
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{overview.png}
+  \caption{Overview of proposed framework.
+Figure on the left shows the setup for clothing assistance and the figure on the right shows the latent space encoding the motor skills.
+This latent space can be used a tool for Learning from Demonstration (LfD) and also for data-efficient reinforcement learning.}
+  \label{figure:overview}
+\end{figure}
 In this study, we explore the suitability of using Bayesian Gaussian Process Latent Variable Model (BGPLVM) to learn a latent space as a low dimensional representation of motor skills for clothing assistance.
-This section is divided into two parts.
-Section~\ref{section:bgplvm} provides the formulation of BGPLVM and Sections~\ref{section:clothassist}, ref{section:controller} presents the application of BGPLVM in the clothing assistance framework.
+This section is divided into three parts.
+Section~\ref{section:bgplvm} provides the formulation of BGPLVM, Section~\ref{section:clothassist} presents the application of BGPLVM in the clothing assistance framework and Section~\ref{section:latentrl} includes the formulation of a policy search framework that relies on the use of this representation.
 
@@
 \subsection{Motor Skills Representation using BGPLVM}
-\label{section:controller}
+\label{section:clothassist}
 
-Motor skills for clothing assistance are given by high dimensional joint angle trajectories of a 14 DOF dual-arm robot.
-The robot also has to maintain several task space constraints such as coupling with a clothing article along with safe human-robot interaction as shown in Figure~\ref{figure:setup}.
+Motor skills for clothing assistance are given by high dimensional joint angle trajectories of the dual-arm robot.
+The robot also has to maintain several task space constraints such as coupling with a clothing article along with safe human-robot interaction as shown in Figure~\ref{figure:overview}.
 To address these problems, we propose the use of BGPLVM for learning a low-dimensional latent space through a nonlinear mapping to the kinematic space.
@@
 
-The dataset, $\mathbf{Y}$, was obtained from successful clothing assistance demonstrations for various postures of the mannequin.
 There can be several types of failure scenarios when the robot performs clothing tasks.
@@
 The BGPLVM model learns a mapping from the low dimensional latent space to the robot kinematic space such that a trajectory of latent points generates a trajectory on the dual-arm robot.
-This interface can be used as a tool for Learning from Demonstration (LfD) where the necessary clothing skills are imparted to the robot by using cursor control over the latent space as shown in Figure~\ref{figure:realtime}.
+This interface can be used as a tool for Learning from Demonstration (LfD) where the necessary clothing skills are imparted to the robot by using cursor control over the latent space as shown in Figure~\ref{figure:overview}.
 
 \subsection{Latent Space Reinforcement Learning}
-\label{section:latentspace}
+\label{section:latentrl}
 
-The latent space learned using BGPLVM is an efficient search space to learn policies that generalize well to unseen environmental settings.
+In this study, we hypothesize that the latent space learned using BGPLVM is an efficient search space to learn policies that generalize well to unseen environmental settings.
+We formulate a policy search framework in the BGPLVM latent space to demonstrate the validity of this hypothesis.
+The objective is to learn a high-dimensional robot trajectory for performing the clothing task to an unseen posture of the mannequin by searching within this latent space.
+Firstly, a dataset of successful clothing assistance trajectories is used to train a latent space that encodes the motor skills.
+Each of the trajectory is now transformed into a sequence of points in the latent space forming latent space trajectories.
+The policy search is performed using the derivative-free Cross Entropy Method which treats the cost as black box function.
+
+We consider Dynamic Movement Primitives (DMP)~\cite{dmp} as the policy representation, commonly used in Robotics.
+It is combination of a point attractor dynamical system and a non-linear forcing term learned using Locally Weighted Regression (LWR)~\cite{lwr}.
+We train a DMP on one of the latent trajectories and the LWR wight coefficients are used as the policy parameters.
+The cost function for policy improvement is designed in the high-dimensional action space.
+In the current setting, we obtain a demonstration for the unseen posture and consider this as the optimized trajectory to be achieved.
+The optimized trajectory is efficiently encoded by via-points extracted using a minimum jerk criterion~\cite{via} that the robot needs to pass through.
+The cost function is given by the sum of all errors between the reconstructed robot trajectory and the desired via-points.
 
@@
 
-In this section, we present the results of using BGPLVM as a representation for clothing assistance motor skills.
-The experimental setup includes the Baxter 14 DOF dual-arm robot performing clothing tasks and a soft mannequin as the subject as shown in Figure~\ref{figure:setup}.
-The clothing task is to cloth the mannequin with a T-shirt which is initially resting on the mannequin's arms.
-The evaluation is presented in two parts: Section~\ref{section:performance} demonstrates the generalization capability of BGPLVM to unseen environmental settings and Section~\ref{section:controllerres} presents the performance of the real-time controller presented in Section~\ref{section:controller}.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.3\textwidth]{latentscales.png}
+  \includegraphics[width=0.3\textwidth]{latentspace.png}
+  \includegraphics[width=0.3\textwidth]{rlcurve.png}
+  \caption{Experimental results of applying BGPLVM to clothing assistance demonstrations for the joint angle scenario.
+Figure on left shows the resultant ARD weight parameters, figure in the middle shows the latent space and figure on the right shows the learning curves comparing reinforcement learning in latent space and action space.}
+  \label{figure:results}
+\end{figure}
 
-\subsection{Controller Demonstration}
-\label{section:controllerres}
+In this section, we present the applicability of BGPLVM for encoding clothing skills.
+We used an evaluation dataset given by successful clothing assistance demonstrations for 4 different postures of the mannequin obtained by changing the elevation of the mannequin's arms.
+Latent spaces were trained by considering 3 trajectories as the training data and the remaining trajectory was used as test data.
+The ARD weights on training resulted in 3 active latent dimensions for the joint angle scenario and 2 latent dimensions for the end effector scenario.
+This implies that the motor skills encoded varies depending the input action space.
+The latent scales and latent space visualizations for the joint angle scenario is shown in Figure~\ref{figure:results}.
 
-In this section, the latent space controller presented in Section~\ref{section:controller} is demonstrated.
-The latent space for the controller is trained for the joint angles of a clothing trial obtained from kinesthetic demonstration by a human.
-The resultant model represented the data in a 2 dimensional latent space.
-Cursor control on the latent space was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points as shown in Figure~\ref{figure:controller}.
+Cursor control on the latent space was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points.
 The resultant latent dimensions further captured a specific aspect of the clothing motor-skills.
@@
 The second dimension captured various vertical motions of pulling up the T-shirt in the beginning and pulling it down along the torso at the end.
+A video demonstration of the controller is available at (youtube link).
+
+We further demonstrate the effectivity of performing policy search in the BGPLVM latent space by comparison with search in the high-dimensional action space.
+For the comparison, the policy search method, policy representation and reward function are the same for both the cases.
+The main difference is the search space i.e.
+low-dimensional latent space or high-dimensional action space.
+CEM was run for 50 iterations where 50 rollouts were performed in each iteration and the top 5 rollouts were retained for computing the mean and variance of the policy parameters.
+Figure~\ref{figure:results} show the learning curves for both reinforcement learning formulations indicating the latent space scenario outperforms policy search in the high dimensional space, thereby validating our hypothesis.
 
@@
 The experimental results indicate our method as a promising approach in combination with reinforcement learning.
+There can be several extensions to this work.
 A latent space representation can be learned from multiple observation spaces using Manifold Relevance Determination (MRD)~\cite{mrd} which is an extension of BGPLVM.
 Based on this flexibility, our future work will be to learn models that explicitly incorporates visual information of the relationship between the human and cloth.
-The long term goal will be to develop a policy search reinforcement learning framework that relies on the use of BGPLVM.
+The long term goal is to develop a data-efficient policy search reinforcement learning framework that unifies learning the latent space simultaneously with policy learning.
 
@@
 
+\bibitem{dmp}
+Ijspeert, A.
+J., Nakanishi, J., Hoffmann, H., Pastor, P., and Schaal, S., ``Dynamical movement primitives: learning attractor models for motor behaviors'', in Neural computation 25, no.
+2 (2013): 328-373.
+
+\bibitem{lwr}
+Schaal, S., and Atkeson, C. G., ``Constructive incremental learning from only local information'', in Neural computation 10.8 (1998): 2047-2084.
+
+\bibitem{via}
+Wada, Y., and Kawato, M., ``A theory for cursive handwriting based on the minimization principle'', in Biological Cybernetics 73.1 (1995): 3-13.
+
 \bibitem{mrd}

### 2016-11-10 Experiments/plotFuncs.py
[minor] inference functions for pca and bgplvm

@@ -110,6 +110,38 @@ def plotTraj2(Dataset, points = None, colors={'Train':'b','Test':'r'}):
     # show all the plots
     plt.show()
 
+def pcaLatent(model, trainInput, testInput):
+    # get latent space plot parameters
+    testData = model.transform(testInput)
+    trainData = model.transform(trainInput)
+    return trainData, testData
+
+def gplvmLatent(model, trainInput, testInput):
+    # get latent space plot parameters
+    qDim = model.X.shape[1]
+    testData = np.zeros((testInput.shape[0], qDim))
+    trainData = np.zeros((trainInput.shape[0], qDim))
+
+    for n in range(trainInput.shape[0]):
+        # infer latent position
+        xTrain, _ = model.infer_newX(np.atleast_2d(trainInput[n,:]), optimize=True)
+
+        # update parameter
+        trainData[n,:] = xTrain
+        sys.stdout.write('.')
+    sys.stdout.write('\n')
+
+    for n in range(testInput.shape[0]):
+        # infer latent position
+        xTest, _ = model.infer_newX(np.atleast_2d(testInput[n,:]), optimize=True)
+
+        # update parameter
+        testData[n,:] = xTest
+        sys.stdout.write('.')
+    sys.stdout.write('\n')
+
+    return trainData, testData
+
 def plotLatent(model, trainInput, testInput, nPoints=400, wThresh=0.05):
     sTest = 200
     sTrain = 150

### 2016-11-10 Paper/paper.tex
[major] second draft for paper

@@
 
-\title{Data-efficient Motor-skills Learning in Latent Spaces for Robotic Clothing Assistance}
+\title{Bayesian Nonparametric Motor-skill Representations \\ for Efficient Learning of Robotic Clothing Assistance}
 
@@
 \begin{abstract}
-Robotic implementation of clothing assistance is a challenging problem that can greatly improve the quality-of-life for the elderly and disabled.
+Motor-skill learning for complex robotic tasks is a challenging problem due to the high task variability.
+Robotic clothing assistance is one such challenging problem that can greatly improve the quality-of-life for the elderly and disabled.
 In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent manifold learning.
@@
 
-Clothing assistance is a necessity in the daily life of the elderly and disabled people.
-However, robotic clothing assistance is still considered an open problem.
-Design of a robust framework involves reliable cloth state estimation in real-time and a motor skills learning framework that can detect and adapt to various failure scenarios.
-Tamei \emph{et al.}~\cite{tamei} proposed a reinforcement learning (RL) framework for clothing assistance with a dual-arm robot to cloth a soft mannequin with a T-shirt initially resting on its arms.
-For tractable learning time, policy update was done using finite difference policy gradient applied to a single via-point of a single joint in each robot arm.
-This severely constrained the generalization capability to very different environmental settings such as major changes in the subject's posture or using different clothing article.
+Enabling robots to interact with the environment and learning motor skills to perform complex tasks is one of the major goals of the AI community.
+However, existing methods require large number of interactions to learn optimal behavior which is not suitable in practical scenarios such as in the field of Assistive Robotics.
+Clothing assistance is one such task which is a necessity in the daily life of the elderly and disabled people.
+Design of a robotic framework involves a motor skills learning framework that can detect and adapt to various failure scenarios.
+Tamei \emph{et al.}~\cite{tamei} proposed a reinforcement learning (RL) framework for clothing assistance with a dual-arm robot to cloth a soft mannequin with a T-shirt.
+For tractable learning time, policy update was done using finite difference policy gradient applied to a small subset of policy parameters.
+This constrained generalization capability to very environmental settings such as major changes in the subject's posture or using different clothing article.
 
@@
 In this study, we propose an efficient representation of motor skills that relies on the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM)~\cite{bgplvm}.
-BGPLVM is capable of learning a data-efficient latent space for clothing tasks performed by a dual-arm robot.
-Representation of clothing skills in a low-dimensional space enables the use of expressive policy update rules for generalization to very different settings.
+We hypothesize that the latent space learned using BGPLVM is an efficient search space to learn novel motor-skills for complex robotics tasks such as clothing assistance.
+Furthermore, the mapping from latent space to data space is modeled using a Gaussian Process (GP)~\cite{gp} allowing data-efficient learning.
+This is crucial to ensure the robot can generalize well to unseen environmental settings within few interactions making it suitable for a feasible social implementation of robotic clothing assistance.
 We apply our proposed method to a clothing assistance setup as shown in Figure~\ref{figure:overview}.
@@
 We further present the design of a real-time controller from the BGPLVM latent space that can be used as a tool for Learning from Demonstration (LfD).
-The experimental results indicate a promising reinforcement learning learning that can be used for challenging tasks such as robotic clothing assistance.
-
-The rest of the paper is structured as follows.
-The proposed framework is presented in Section~\ref{section:method}.
-Section~\ref{section:results} includes experimental results and Section~\ref{section:conclusion} concludes the paper with directions for future work.
+The experimental results indicate a promising reinforcement learning framework that can be used for challenging tasks such as robotic clothing assistance.
 
@@
 \end{figure}
-In this study, we explore the suitability of using Bayesian Gaussian Process Latent Variable Model (BGPLVM) to learn a latent space as a low dimensional representation of motor skills for clothing assistance.
-This section is divided into three parts.
-Section~\ref{section:bgplvm} provides the formulation of BGPLVM, Section~\ref{section:clothassist} presents the application of BGPLVM in the clothing assistance framework and Section~\ref{section:latentrl} includes the formulation of a policy search framework that relies on the use of this representation.
+Motor skills for clothing assistance are given by high dimensional joint angle trajectories of the dual-arm robot.
+The robot also has to maintain several task space constraints such as coupling with a clothing article along with safe human-robot interaction as shown in Figure~\ref{figure:overview}.
+To address these problems, we propose the use of BGPLVM for learning a low-dimensional latent space through a nonlinear mapping to the kinematic space.
+There several motivating factors for this choice.
+The Bayesian treatment avoids over fitting to the training data and the use of an ARD kernel in the GP mapping leads to inference of the inherent dimensionality to encode motor-skills.
 
@@
 \end{equation}
-Efficient computations to handle test data is further described in \cite{bgplvm}.
 
@@
 
-Motor skills for clothing assistance are given by high dimensional joint angle trajectories of the dual-arm robot.
-The robot also has to maintain several task space constraints such as coupling with a clothing article along with safe human-robot interaction as shown in Figure~\ref{figure:overview}.
-To address these problems, we propose the use of BGPLVM for learning a low-dimensional latent space through a nonlinear mapping to the kinematic space.
-The Bayesian treatment avoids over fitting to the training data and automatic inference of latent space dimensionality through the ARD kernel.
+In this section, we present the formulation used to apply BGPLVM to clothing assistance skills.
 We consider the clothing task where a dual-arm robot clothes a soft mannequin with a T-shirt which is initially resting on the mannequin's arms.
@@
 There can be several types of failure scenarios when the robot performs clothing tasks.
-While clothing, the T-shirt collar could get stuck on the mannequin's head and while unclothing, the T-shirt could get stuck in the mannequin's shoulder joints.
 To recover from these failures, not only is the trajectory of the robot important, but also the speed of execution.
@@
 The BGPLVM model learns a mapping from the low dimensional latent space to the robot kinematic space such that a trajectory of latent points generates a trajectory on the dual-arm robot.
-This interface can be used as a tool for Learning from Demonstration (LfD) where the necessary clothing skills are imparted to the robot by using cursor control over the latent space as shown in Figure~\ref{figure:overview}.
+This interface can be used as a tool for Learning from Demonstration (LfD) where the necessary clothing skills are imparted to the robot by using cursor control over the latent space.
+Cursor control on the latent space was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points.
 
@@
 
-In this study, we hypothesize that the latent space learned using BGPLVM is an efficient search space to learn policies that generalize well to unseen environmental settings.
-We formulate a policy search framework in the BGPLVM latent space to demonstrate the validity of this hypothesis.
-The objective is to learn a high-dimensional robot trajectory for performing the clothing task to an unseen posture of the mannequin by searching within this latent space.
+In this section, we formulate a policy search framework in the BGPLVM latent space.
+The objective is to learn a high-dimensional robot trajectory for performing the clothing task on an unseen posture of the mannequin by searching within this latent space.
 Firstly, a dataset of successful clothing assistance trajectories is used to train a latent space that encodes the motor skills.
 Each of the trajectory is now transformed into a sequence of points in the latent space forming latent space trajectories.
-The policy search is performed using the derivative-free Cross Entropy Method which treats the cost as black box function.
+The policy search is performed using the cross-entropy method (CEM)~\cite{cem} which is a general algorithm for approximately solving global optimization problems.
+It has been widely used in robotics to learn controllers for performing various tasks.
 
-We consider Dynamic Movement Primitives (DMP)~\cite{dmp} as the policy representation, commonly used in Robotics.
+We consider Dynamic Movement Primitives (DMP)~\cite{dmp} as the policy representation.
 It is combination of a point attractor dynamical system and a non-linear forcing term learned using Locally Weighted Regression (LWR)~\cite{lwr}.
+%\begin{equation}
+%  \begin{array}{c}
+%    \tau \ddot{x} = K(g-x) -D\dot{x} + (g - x_0)f\\~\\
+%    \displaystyle f(s) = \frac{\sum_i w_i \psi_i(s)s}{\sum_i \psi_i(s)}, \text{ where } \tau \dot{s} = - \alpha s
+%  \end{array}
+%\end{equation}
 We train a DMP on one of the latent trajectories and the LWR wight coefficients are used as the policy parameters.
@@
 In the current setting, we obtain a demonstration for the unseen posture and consider this as the optimized trajectory to be achieved.
-The optimized trajectory is efficiently encoded by via-points extracted using a minimum jerk criterion~\cite{via} that the robot needs to pass through.
-The cost function is given by the sum of all errors between the reconstructed robot trajectory and the desired via-points.
-
+The optimized trajectory is efficiently encoded by via-points extracted using the minimum jerk criterion~\cite{via} that the robot needs to pass through.
+The cost function is given by the sum of all errors between the reconstructed robot trajectory and the desired via-points:
+\begin{equation}
+  R(\tau) = \sum_{i = 1}^{n_{\text{dims}}} \sum_{j = 1}^{n_{\text{via}}} (V_{i,j} - x^{recons}_i(t_{i,j}))
+\end{equation}
+where $R(\tau)$ is the total reward for trajectory $\tau$, $V_{i,j}$ is the $j^{\text{th}}$ via point of $i^{\text{th}}$ dimension and $x^{recons}_i(t)$ is the value at time $t$ for $i^{\text{th}}$ dimension of reconstructed trajectory.
 \section{Results}
@@
   \centering
-  \includegraphics[width=0.3\textwidth]{latentscales.png}
   \includegraphics[width=0.3\textwidth]{latentspace.png}
+  \includegraphics[width=0.3\textwidth]{rmse.png}
   \includegraphics[width=0.3\textwidth]{rlcurve.png}
   \caption{Experimental results of applying BGPLVM to clothing assistance demonstrations for the joint angle scenario.
-Figure on left shows the resultant ARD weight parameters, figure in the middle shows the latent space and figure on the right shows the learning curves comparing reinforcement learning in latent space and action space.}
+Figure on the left shows the latent space, middle shows comparison of reconstruction error and right shows learning curves with different LVMs and in the action space.}
   \label{figure:results}
@@
 In this section, we present the applicability of BGPLVM for encoding clothing skills.
-We used an evaluation dataset given by successful clothing assistance demonstrations for 4 different postures of the mannequin obtained by changing the elevation of the mannequin's arms.
-Latent spaces were trained by considering 3 trajectories as the training data and the remaining trajectory was used as test data.
+The evaluation dataset contains 4 clothing trials that were obtained for 4 different postures of the mannequin wherein the elevation of the arms were varied forming different angles $\{65^o, 70^o, 75^o, 80^o\}$ with the mannequin's body.
+The clothing trial for $75^o$ was left out as the test data and latent variable models were trained for the remaining 3 clothing trials.
 The ARD weights on training resulted in 3 active latent dimensions for the joint angle scenario and 2 latent dimensions for the end effector scenario.
@@
 
-Cursor control on the latent space was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points.
-The resultant latent dimensions further captured a specific aspect of the clothing motor-skills.
-For example, the most significant dimension captured the horizontal motion of the arms along the mannequin while maintaining the constraints for clothing.
+The latent dimensions captured by BGPLVM a specific aspect of the clothing motor-skills.
+For the joint angle scenario, the most significant dimension captured the horizontal motion of the arms along the mannequin while maintaining the constraints for clothing.
 The second dimension captured various vertical motions of pulling up the T-shirt in the beginning and pulling it down along the torso at the end.
-A video demonstration of the controller is available at (youtube link).
+The third dimension captured variations in joint configurations across the demonstrations and could explain the constraint of safe human-robot interaction.
+A video demonstration with the exploration of the latent dimensions is available at (youtube link).
+The predictive performance of BGPLVM was evaluated by comparing the reconstruction error with two other latent variable models (LVM) i.e.
+Principal Component Analysis (PCA)~\cite{pca} and Gaussian Process Latent Variable Model (GPLVM)~\cite{gplvm}.
+Figure~\ref{figure:results} shows that BGPLVM has the best performance.
 
-We further demonstrate the effectivity of performing policy search in the BGPLVM latent space by comparison with search in the high-dimensional action space.
+We further demonstrate the effectivity of performing policy search in the BGPLVM latent space by comparison with search in the high-dimensional action space and in latent spaces learned using PCA and GPLVM.
 For the comparison, the policy search method, policy representation and reward function are the same for both the cases.
@@
 CEM was run for 50 iterations where 50 rollouts were performed in each iteration and the top 5 rollouts were retained for computing the mean and variance of the policy parameters.
-Figure~\ref{figure:results} show the learning curves for both reinforcement learning formulations indicating the latent space scenario outperforms policy search in the high dimensional space, thereby validating our hypothesis.
+Figure~\ref{figure:results} show the learning curves for both reinforcement learning formulations indicating the BGPLVM scenario performs better than all other search spaces.
+This validates our hypothesis that encoding motor-skills using Bayesian non parametric latent variable models is an effective parametrization for learning.
 
@@
 
+\bibitem{gp}
+Williams, C., and Rasmussen, C. E., ``Gaussian processes for machine learning'' by the MIT Press 2.3 (2006): 4.
+
+\bibitem{cem}
+De Boer, P., Kroese, D. P., Mannor, S., and Rubinstein, R. Y., ``A tutorial on the cross-entropy method'' in Annals of operations research 134, no. 1 (2005): 19-67.
+
 \bibitem{dmp}
@@
 
+\bibitem{pca}
+Jolliffe, I., ``Principal component analysis'' by John Wiley and Sons, Ltd, 2002.
+
+\bibitem{gplvm}
+Lawrence, N., ``Probabilistic non-linear principal component analysis with Gaussian process latent variable models'' in Journal of Machine Learning Research 6.Nov (2005): 1783-1816.
+
 \bibitem{mrd}

### 2016-11-11 Paper/paper.tex
[major] complete third draft of the paper

@@
   \centering
-  \includegraphics[width=0.8\textwidth]{overview.png}
+  \includegraphics[width=0.8\textwidth, height=0.2\textheight]{overview.png}
   \caption{Overview of proposed framework.
@@
 We represent the robot poses in two ways i.e.
-1) Kinematic representation given by joint angles $D_K = 14$ and 2) Task space representation given by the end-effector pose of both arms with cartesian position ${P_X,P_Y,P_Z} \in \mathbb{R}^3$ and orientation ${O_X,O_Y,O_Z,O_{\omega}} \in \mathbb{R}^4$ forming a 14-dimensional space $D_T = 14$.
+1) Kinematic representation given by joint angles $D_K = 14$ and 2) Task space representation given by the end-effector pose of both arms with Cartesian position ${P_X,P_Y,P_Z} \in \mathbb{R}^3$ and orientation ${O_X,O_Y,O_Z,O_{\omega}} \in \mathbb{R}^4$ forming a 14-dimensional space $D_T = 14$.
 We set the dimension of the latent space as $q = 5$, however the dimensionality is eventually inferred through the training of the ARD kernel weights as explained in Section~\ref{section:bgplvm}.
@@
 This interface can be used as a tool for Learning from Demonstration (LfD) where the necessary clothing skills are imparted to the robot by using cursor control over the latent space.
-Cursor control on the latent space was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points.
+The controller was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points as shown in the video demonstration available at \url{https://youtu.be/2TCZnt_qBHU}.
 
@@
   \centering
-  \includegraphics[width=0.3\textwidth]{latentspace.png}
-  \includegraphics[width=0.3\textwidth]{rmse.png}
-  \includegraphics[width=0.3\textwidth]{rlcurve.png}
-  \caption{Experimental results of applying BGPLVM to clothing assistance demonstrations for the joint angle scenario.
-Figure on the left shows the latent space, middle shows comparison of reconstruction error and right shows learning curves with different LVMs and in the action space.}
+  \subfloat[][Latent Space]{\label{figure:latentspace}
+  \includegraphics[width=0.3\textwidth,height=0.13\textheight]{latentspace.png}}
+  \subfloat[][Reconstruction Error]{\label{figure:rmse}
+  \includegraphics[width=0.3\textwidth,height=0.13\textheight]{rmse.png}}
+  \subfloat[][Learning Curves for RL]{\label{figure:rlcurve}
+  \includegraphics[width=0.3\textwidth,height=0.13\textheight]{rlcurve.png}}
+  \caption{Experimental results of applying BGPLVM to clothing assistance demonstrations for the joint angle scenario.}
   \label{figure:results}
@@
 This implies that the motor skills encoded varies depending the input action space.
-The latent scales and latent space visualizations for the joint angle scenario is shown in Figure~\ref{figure:results}.
+The latent space learned for the joint angle scenario is shown in Figure~\ref{figure:latentspace}.
 
@@
 The third dimension captured variations in joint configurations across the demonstrations and could explain the constraint of safe human-robot interaction.
-A video demonstration with the exploration of the latent dimensions is available at (youtube link).
+A video demonstration with the exploration of the latent dimensions is available at \url{https://youtu.be/2TCZnt_qBHU}.
 The predictive performance of BGPLVM was evaluated by comparing the reconstruction error with two other latent variable models (LVM) i.e.
 Principal Component Analysis (PCA)~\cite{pca} and Gaussian Process Latent Variable Model (GPLVM)~\cite{gplvm}.
-Figure~\ref{figure:results} shows that BGPLVM has the best performance.
+Figure~\ref{figure:rmse} shows that BGPLVM has the best performance.
 
 We further demonstrate the effectivity of performing policy search in the BGPLVM latent space by comparison with search in the high-dimensional action space and in latent spaces learned using PCA and GPLVM.
-For the comparison, the policy search method, policy representation and reward function are the same for both the cases.
-The main difference is the search space i.e.
-low-dimensional latent space or high-dimensional action space.
+For the comparison, the policy search method, policy representation and reward function are the same and the main difference is the search space.
 CEM was run for 50 iterations where 50 rollouts were performed in each iteration and the top 5 rollouts were retained for computing the mean and variance of the policy parameters.
-Figure~\ref{figure:results} show the learning curves for both reinforcement learning formulations indicating the BGPLVM scenario performs better than all other search spaces.
-This validates our hypothesis that encoding motor-skills using Bayesian non parametric latent variable models is an effective parametrization for learning.
+Figure~\ref{figure:rlcurve} shows the learning curves for all formulations and the BGPLVM scenario performs better than all other search spaces.
+This validates our hypothesis that encoding motor-skills using Bayesian nonparametric latent variable models is an effective parametrization for learning.
 
@@
 In this study, we have presented the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM) as a representation for encoding motor-skills to perform clothing assistance task.
-The experimental results indicate our method as a promising approach in combination with reinforcement learning.
+The experimental results indicate our method as a promising approach for learning in combination with reinforcement learning.
+We further demonstrated its applicability as an intuitive and user-friendly tool for Learning from Demonstration (LfD).
 There can be several extensions to this work.
-A latent space representation can be learned from multiple observation spaces using Manifold Relevance Determination (MRD)~\cite{mrd} which is an extension of BGPLVM.
-Based on this flexibility, our future work will be to learn models that explicitly incorporates visual information of the relationship between the human and cloth.
+A latent space representation can be learned from multiple observation spaces using Manifold Relevance Determination (MRD)~\cite{mrd}.
+Our future work will be to learn models that incorporates visual information of the relationship between the human and cloth.
 The long term goal is to develop a data-efficient policy search reinforcement learning framework that unifies learning the latent space simultaneously with policy learning.

### 2016-11-27 Paper/paper.tex
[major] complete 3rd draft of paper
[minor] edit minor comments for camera ready paper

@@
 Robotic clothing assistance is one such challenging problem that can greatly improve the quality-of-life for the elderly and disabled.
-In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent manifold learning.
-We implement our framework in a practical setting with a dual-arm robot performing clothing tasks.
+In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent space learning.
+We implement our framework in a practical setting with a dual-arm robot performing clothing assistance tasks.
 We demonstrate that performing motor-skills learning in such task specific latent spaces outperforms learning in the high-dimensional joint configuration space of the robot.
@@
 For tractable learning time, policy update was done using finite difference policy gradient applied to a small subset of policy parameters.
-This constrained generalization capability to very environmental settings such as major changes in the subject's posture or using different clothing article.
+This constrained the generalization capability of the algorithm to very different environmental settings such as major changes in the subject's posture or using different clothing articles.
 
@@
 An efficient approach is the use of dimensionality reduction (DR) along with reinforcement learning (RL).
-There have been several studies in this direction in the recent years.
-Bitzer \emph{et al.}~\cite{bitzer} proposed a learning framework where DR was used as a preprocessing step to extract the problem structure.
+There have been several studies in this direction in recent years.
+Bitzer \emph{et al.}~\cite{bitzer} proposed a learning framework where DR was used as a preprocessing step to extract problem structure.
 However, this study relies on a maximum-a-posteriori (MAP) estimate of the latent space and can overfit to the training data.
@@
 In this study, we propose an efficient representation of motor skills that relies on the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM)~\cite{bgplvm}.
-We hypothesize that the latent space learned using BGPLVM is an efficient search space to learn novel motor-skills for complex robotics tasks such as clothing assistance.
-Furthermore, the mapping from latent space to data space is modeled using a Gaussian Process (GP)~\cite{gp} allowing data-efficient learning.
+We hypothesize that the latent space learned using BGPLVM is an efficient search space to learn novel motor-skills for complex robotic tasks such as clothing assistance.
+The mapping from latent space to data space is modeled using a Gaussian Process (GP)~\cite{gp} allowing data-efficient learning.
 This is crucial to ensure the robot can generalize well to unseen environmental settings within few interactions making it suitable for a feasible social implementation of robotic clothing assistance.
@@
 Figure on the left shows the setup for clothing assistance and the figure on the right shows the latent space encoding the motor skills.
-This latent space can be used a tool for Learning from Demonstration (LfD) and also for data-efficient reinforcement learning.}
+This latent space can be used as a tool for Learning from Demonstration (LfD) and also for data-efficient reinforcement learning.}
   \label{figure:overview}
@@
 To address these problems, we propose the use of BGPLVM for learning a low-dimensional latent space through a nonlinear mapping to the kinematic space.
-There several motivating factors for this choice.
-The Bayesian treatment avoids over fitting to the training data and the use of an ARD kernel in the GP mapping leads to inference of the inherent dimensionality to encode motor-skills.
+There are several motivating factors for this choice.
+The Bayesian treatment avoids overfitting to the training data and the use of an ARD kernel in the GP mapping leads to inference of the inherent dimensionality to encode motor-skills.
 
@@
 \end{equation}
-where $\mathbf{X},\Phi$ are the unknown latent points and hyper parameters for the GP mapping that need to be inferred.
+where $\mathbf{X},\Phi$ are the unknown latent points and hyperparameters for the GP mapping that need to be inferred.
 $\mathbf{K}$ is the kernel matrix constructed from the latent points.
@@
 
-We consider Dynamic Movement Primitives (DMP)~\cite{dmp} as the policy representation.
-It is combination of a point attractor dynamical system and a non-linear forcing term learned using Locally Weighted Regression (LWR)~\cite{lwr}.
+We consider Dynamic Movement Primitives (DMP)~\cite{dmp} as policy representation.
+It is the combination of a point attractor dynamical system and a non-linear forcing term learned using Locally Weighted Regression (LWR)~\cite{lwr}.
 %\begin{equation}
@@
 %\end{equation}
-We train a DMP on one of the latent trajectories and the LWR wight coefficients are used as the policy parameters.
+We train a DMP on one of the latent trajectories and the LWR weight coefficients are used as policy parameters.
 The cost function for policy improvement is designed in the high-dimensional action space.
@@
 The ARD weights on training resulted in 3 active latent dimensions for the joint angle scenario and 2 latent dimensions for the end effector scenario.
-This implies that the motor skills encoded varies depending the input action space.
+This implies that the encoded motor skills varies depending on the input action space.
 The latent space learned for the joint angle scenario is shown in Figure~\ref{figure:latentspace}.
@@
 We further demonstrate the effectivity of performing policy search in the BGPLVM latent space by comparison with search in the high-dimensional action space and in latent spaces learned using PCA and GPLVM.
-For the comparison, the policy search method, policy representation and reward function are the same and the main difference is the search space.
-CEM was run for 50 iterations where 50 rollouts were performed in each iteration and the top 5 rollouts were retained for computing the mean and variance of the policy parameters.
+CEM was run for 50 iterations where 50 rollouts were performed in each iteration.
+The rollouts were generated from DMP with random sampling of LWR weight parameters.
+The top 5 rollouts were retained for updating the mean and variance of policy parameters at each iteration.
 Figure~\ref{figure:rlcurve} shows the learning curves for all formulations and the BGPLVM scenario performs better than all other search spaces.
+PCA suffers from linearity assumptions and as GPLVM relies on a MAP estimate of the latent space it could be overfitting to the training data.
 This validates our hypothesis that encoding motor-skills using Bayesian nonparametric latent variable models is an effective parametrization for learning.

### 2016-12-05 Paper/paper.tex
[minor] complete changes for camera ready draft

@@
 
-Enabling robots to interact with the environment and learning motor skills to perform complex tasks is one of the major goals of the AI community.
+Enabling robots to learn motor skills for performing complex tasks is one of the major goals of the AI community.
 However, existing methods require large number of interactions to learn optimal behavior which is not suitable in practical scenarios such as in the field of Assistive Robotics.
@@
 
-To address this problem, we need to consider an alternate representation that is more flexible and suitable for robust learning of motor-skills.
-An efficient approach is the use of dimensionality reduction (DR) along with reinforcement learning (RL).
-There have been several studies in this direction in recent years.
-Bitzer \emph{et al.}~\cite{bitzer} proposed a learning framework where DR was used as a preprocessing step to extract problem structure.
-However, this study relies on a maximum-a-posteriori (MAP) estimate of the latent space and can overfit to the training data.
-Luck \emph{et al.}~\cite{luck} proposed a policy search framework that inherently combines RL and DR wherein the latent space is learned based on the reward signals.
-However, this framework relies on the use of linear dimensionality reduction which limits applicability to complex tasks.
-
+To address this problem, we consider an alternate representation that is more flexible and suitable for robust learning of motor-skills.
+An efficient approach is the use of dimensionality reduction (DR) along with reinforcement learning (RL)~\cite{bitzer}, \cite{luck}.
 In this study, we propose an efficient representation of motor skills that relies on the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM)~\cite{bgplvm}.
-We hypothesize that the latent space learned using BGPLVM is an efficient search space to learn novel motor-skills for complex robotic tasks such as clothing assistance.
+We hypothesize that the latent space learned using BGPLVM is an efficient search space to learn novel motor-skills.
 The mapping from latent space to data space is modeled using a Gaussian Process (GP)~\cite{gp} allowing data-efficient learning.
@@
   \centering
-  \includegraphics[width=0.8\textwidth, height=0.2\textheight]{overview.png}
-  \caption{Overview of proposed framework.
-Figure on the left shows the setup for clothing assistance and the figure on the right shows the latent space encoding the motor skills.
-This latent space can be used as a tool for Learning from Demonstration (LfD) and also for data-efficient reinforcement learning.}
+  \includegraphics[width=0.76\textwidth, height=0.16\textheight]{overview.png}
+  \caption{Overview of proposed framework. Figure on the left shows the setup for clothing assistance and the figure on the right shows the latent space encoding the motor skills.}
   \label{figure:overview}
@@
 The mapping function $f$ is modeled using a Gaussian Process (GP) suitable for performing non-linear dimensionality reduction with use of a non-linear kernel function.
-The marginal likelihood for the generative model is given by:
+The conditional likelihood for the generative model is given by:
 \begin{equation}
-  \label{eqn:marginallikelihood}
+  \label{eqn:conditionallikelihood}
   p(\mathbf{Y}|\mathbf{X},\Phi) = \prod_{d=1}^{D} \mathcal{N}(\mathbf{y}_{:,d}|\mathbf{0},\mathbf{K}+\beta^{-1}I)\\
@@
 In the generative model, the latent positions need to be marginalized out for having a purely Bayesian treatment.
-Titsias \emph{et al.}~\cite{bgplvm} proposed a variational inference approach to compute a tractable lower bound for the marginalization thereby inferring a posterior distribution on the latent positions rather than a MAP estimate.
+Titsias \emph{et al.}~\cite{bgplvm} proposed a variational inference approach to compute a tractable lower bound for the marginalization thereby inferring a posterior distribution on the latent space which avoids the problem of overfitting.
 
-For performing automatic model selection of the latent space dimensionality, the Automatic Relevance Determination (ARD) Kernel can be used in the GP mapping:
+For performing automatic model selection of the latent space dimensionality, the Automatic Relevance Determination (ARD) Kernel~\cite{gp} can be used in the GP mapping:
 \begin{equation}
@@
 1) Kinematic representation given by joint angles $D_K = 14$ and 2) Task space representation given by the end-effector pose of both arms with Cartesian position ${P_X,P_Y,P_Z} \in \mathbb{R}^3$ and orientation ${O_X,O_Y,O_Z,O_{\omega}} \in \mathbb{R}^4$ forming a 14-dimensional space $D_T = 14$.
-We set the dimension of the latent space as $q = 5$, however the dimensionality is eventually inferred through the training of the ARD kernel weights as explained in Section~\ref{section:bgplvm}.
+We set the dimension of the latent space as $q = 5$.
+However, the dimensionality is eventually inferred through the training of the ARD kernel weights as explained in Section~\ref{section:bgplvm}.
 
@@
 This interface can be used as a tool for Learning from Demonstration (LfD) where the necessary clothing skills are imparted to the robot by using cursor control over the latent space.
-The controller was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points as shown in the video demonstration available at \url{https://youtu.be/2TCZnt_qBHU}.
+The controller was able to reproduce the clothing demonstration even when the latent trajectory was different from the training latent points.
 
@@
 We consider Dynamic Movement Primitives (DMP)~\cite{dmp} as policy representation.
-It is the combination of a point attractor dynamical system and a non-linear forcing term learned using Locally Weighted Regression (LWR)~\cite{lwr}.
-%\begin{equation}
-%  \begin{array}{c}
-%    \tau \ddot{x} = K(g-x) -D\dot{x} + (g - x_0)f\\~\\
-%    \displaystyle f(s) = \frac{\sum_i w_i \psi_i(s)s}{\sum_i \psi_i(s)}, \text{ where } \tau \dot{s} = - \alpha s
-%  \end{array}
-%\end{equation}
-We train a DMP on one of the latent trajectories and the LWR weight coefficients are used as policy parameters.
+It is the combination of a point attractor dynamical system and a non-linear forcing term ($f(s)$) learned using Locally Weighted Regression (LWR)~\cite{lwr}:
+\begin{equation}
+  \begin{array}{c}
+    \tau \ddot{x} = K(g-x) -D\dot{x} + (g - x_0)f\\
+    \displaystyle f(s) = \frac{\sum_i w_i \psi_i(s)s}{\sum_i \psi_i(s)}, \text{ where } \tau \dot{s} = - \alpha s
+  \end{array}
+\end{equation}
+We train a DMP on the latent points corresponding to one of the training trajectory and the LWR weight coefficients are used as policy parameters which are modified to generalize to unseen environmental settings.
 The cost function for policy improvement is designed in the high-dimensional action space.
@@
 \begin{equation}
-  R(\tau) = \sum_{i = 1}^{n_{\text{dims}}} \sum_{j = 1}^{n_{\text{via}}} (V_{i,j} - x^{recons}_i(t_{i,j}))
+  R(\tau) = \sum_{i = 1}^{n_{\text{dims}}} \sum_{j = 1}^{n_{\text{via}}} \| V_{i,j} - x^{recons}_i(t_{i,j}) \|^2
 \end{equation}
@@
 The evaluation dataset contains 4 clothing trials that were obtained for 4 different postures of the mannequin wherein the elevation of the arms were varied forming different angles $\{65^o, 70^o, 75^o, 80^o\}$ with the mannequin's body.
+These arm inclinations cover the entire range for which the robot can successfully perform clothing tasks thereby spanning all feasible postures.
 The clothing trial for $75^o$ was left out as the test data and latent variable models were trained for the remaining 3 clothing trials.
@@
 The latent space learned for the joint angle scenario is shown in Figure~\ref{figure:latentspace}.
+The computational complexity of the proposed framework depends on the forward mapping from the BGPLVM latent space to joint angle space which scales with the size of the training dataset and the number of inducing points used to approximate the dataset i.e.
+$O(nm^2)$~\cite{bgplvm}.
+In the current framework, the training dataset remains fixed during runtime and so precomputated values of the kernel matrix can be used to reduce time-complexity to $O(n)$.
 
@@
 A video demonstration with the exploration of the latent dimensions is available at \url{https://youtu.be/2TCZnt_qBHU}.
-The predictive performance of BGPLVM was evaluated by comparing the reconstruction error with two other latent variable models (LVM) i.e.
+The The predictive performance of BGPLVM was evaluated by comparing the reconstruction error with two other latent variable models (LVM) i.e.
 Principal Component Analysis (PCA)~\cite{pca} and Gaussian Process Latent Variable Model (GPLVM)~\cite{gplvm}.
@@
 Figure~\ref{figure:rlcurve} shows the learning curves for all formulations and the BGPLVM scenario performs better than all other search spaces.
-PCA suffers from linearity assumptions and as GPLVM relies on a MAP estimate of the latent space it could be overfitting to the training data.
+PCA relies on a linear model thereby constraining its capability.
+GPLVM has the worst performance among the latent spaces as relies on a MAP estimate of the latent space and it is overfitting to the training data.
 This validates our hypothesis that encoding motor-skills using Bayesian nonparametric latent variable models is an effective parametrization for learning.
@@
 
-Motor-skills learning for robotic clothing assistance task involves a high dimensional kinematic search space and maintaining several task space constraints.
 In this study, we have presented the use of Bayesian Gaussian Process Latent Variable Model (BGPLVM) as a representation for encoding motor-skills to perform clothing assistance task.

### 2016-12-05 Poster/poster.tex
[minor] add template files for poster

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}

\usepackage[orientation=portrait,size=a0,scale=1.4]{beamerposter}
\usetheme{I6pd2}
\usepackage[english]{babel}
\usepackage{amsmath,amsthm,amssymb,latexsym}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{xcolor}
\graphicspath{{Images/}}
\usepackage[tight,footnotesize]{subfigure}

\definecolor{Blue}{HTML}{000080}

\usecaptiontemplate{\small\structure{\insertcaptionname~\insertcaptionnumber: }\insertcaption}

\title{Bayesian Nonparametric Motor-skill Representations \\ for Efficient Learning of Robotic Clothing Assistance}
\author{Nishanth Koganti$^{1,2}$, Tomoya Tamei$^1$, Kazushi Ikeda$^1$, Tomohiro Shibata$^2$}
\institute{$^1$Nara Institute of Science and Technology, Japan~~$^2$Kyushu Institute of Technology, Japan}

\newcommand{\leftfoot}{Neural Information Processing Systems 2016}
\newcommand{\rightfoot}{Workshop on Practical Bayesian Nonparametrics}

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{0.5ex}}
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}}

\begin{frame}[t]

\begin{columns}[t]

\begin{column}{.02\linewidth}\end{column}

\begin{column}{0.47\linewidth}

\begin{block}{Introduction}

\begin{center}
\textbf{Objective}: Robotic Clothing Assistance is a challenging problem involving \emph{interaction with Human and Clothing material}. In this study, we propose a method for the realtime estimation of the Human-Cloth topological relationship using a depth sensor.
\end{center}

\begin{columns}[t]

\begin{column}{0.55\linewidth}

~\\
Tamei et. al.[1] developed clothing assistance robot to perform T-shirt clothing task:\\~\\
\begin{itemize}
\item \textbf{Reinforcement Learning} scheme used to acquire motor skills for cloth handling.\\~\\
\item \textbf{Low dimensional represetations} of state and policy used to ensure fast learning time.
\end{itemize}

\end{column}

\begin{column}{0.45\linewidth}

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{mannequin.png}
\end{figure}

\end{column}

\end{columns}

\end{block}

\end{column}

\begin{column}{0.47\linewidth}

\begin{block}{Problem Decription}

\begin{columns}[t]

\begin{column}{0.5\linewidth}
\textbf{Topology Coordinates}[2] represents interaction between two curves using low dimensional variables:

\begin{itemize}
\item \textbf{Writhe}, $w \in \mathbb{R}$
\vspace{3mm}
\item \textbf{Center}, $\text{c} = [c_1,c_2] \in \mathbb{R}^2$
\vspace{3mm}
\item \textbf{Density}, $d \in \mathbb{R}$
\end{itemize}

\vspace{3mm}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{humancloth.pdf}
\end{figure}

\end{column}

\begin{column}{0.5\linewidth}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{topology.png}
\end{figure}

\vspace{3mm}

\centering \textbf{Human-Cloth relationship} low dimensionally represented using topology coordinates:

\begin{itemize}
\item T-shirt collar and Mannequin's Head
\vspace{2mm}
\item T-shirt collar and Mannequin's Body
\vspace{2mm}
\item T-shirt sleeve and Mannequin's Shoulder
\end{itemize}

\end{column}

\end{columns}

\end{block}

\end{column}

\begin{column}{.02\linewidth}\end{column}

\end{columns}

\begin{columns}[t]

\begin{column}{0.02\linewidth}\end{column}

\begin{column}{0.96\linewidth}

\begin{alertblock}{Proposed Method}

\begin{columns}[t]

\begin{column}{0.24\linewidth}

\centering \textbf{Experimental Setup}

\begin{itemize}
\item Human-Cloth relationship observed using \textbf{Kinect depth sensor}.\\~\\
\item \textbf{Color markers} used on T-shirt and Mannequin to ensure robust estimation.
\end{itemize}

\vspace{6mm}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{setup.png}
\end{figure}

\end{column}

\begin{column}{0.48\linewidth}

\centering \textbf{Proposed Method}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=0.22\textheight]{method.pdf}
\end{figure}

\end{column}

\begin{column}{0.24\linewidth}

\centering \textbf{T-shirt state estimation}

\vspace{3mm}

\begin{itemize}
\item T-shirt collar and sleeve shape compactly represented using \textbf{Ellipses}.\\~\\
\item Generic 3D conic fitting approach proposed Shakarji et. al.~\cite{ellipse} used for this purpose.
\end{itemize}

\vspace{3mm}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{collarestimates.pdf}
\end{figure}

\end{column}

\end{columns}

\end{alertblock}

\begin{alertblock}{Experimental Results}

\begin{columns}[t]

\begin{column}{0.49\linewidth}

\centering \textbf{Robustness of Topology Space to Measurement Noise}\\
\vspace{3mm}
\centering Robustness evaluated by comparing rewards obtained in \emph{Cartesian} and \emph{Topological} space.

\begin{itemize}
\item Reward function in Cartesian Space:
\begin{equation}
r(o_t) = - \sqrt{(o_t - o_{target})^T \Sigma_o^{-1} (o_t - o_{target})}
\end{equation}
\item Reward function in Topology Space:
\begin{equation}
r(s_t) = - \sqrt{(s_t - s_{target})^T \Sigma_s^{-1} (s_t - s_{target})}
\end{equation}
\end{itemize}

\begin{columns}[t]

\begin{column}{0.49\linewidth}

\centering Simulation Experiment
\begin{figure}
\centering
\includegraphics[width=\textwidth]{toptest.pdf}
\end{figure}

\end{column}

\begin{column}{0.49\linewidth}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{reward.png}
\end{figure}

\end{column}

\end{columns}

\end{column}

\begin{column}{0.49\linewidth}

\centering \textbf{Accuracy of Proposed Method}

\begin{itemize}
\item Accuracy of proposed method evaluated by comparing our estimates with the ground truth.
\vspace{3mm}
\item \textbf{Ground Truth}: Motion Capture System with 7 IR Cameras, 10 IR Markers used on T-shirt collar and 1 IR Marker on T-shirt sleeve.
\end{itemize}

\begin{columns}[t]

\begin{column}{0.59\linewidth}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{realtimeresults.pdf}
\end{figure}

\end{column}

\begin{column}{0.39\linewidth}

\begin{table}
\centering
\begin{center}
\small
\begin{tabular}{ccc}
\toprule
Parameter & Pearson Corr & RMS Error\\[0.2cm]
\midrule
$s_1$ topology & 0.9364 & 0.5955\\
$s_2$ topology & 0.8857 & 0.1247\\
$s_3$ topology & 0.9889 & 0.3529\\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\vspace{4mm}

\begin{itemize}
\item Proposed method provides reliable estimates in real-time (28-30 fps).
\vspace{3mm}
\item However, there are jitters due to inherent noise in depth sensor.
\end{itemize}

\end{column}

\end{columns}

\end{column}

\end{columns}

\end{alertblock}

\end{column}

\begin{column}{0.02\textwidth}\end{column}

\end{columns}

\begin{columns}[t]

\begin{column}{.02\textwidth}\end{column}

\begin{column}{0.47\textwidth}

\begin{block}{Future Work}

\begin{itemize}
\item Bayesian estimation of Human-Cloth relationship using \emph{Proprioceptive} information of robot.
\vspace{3mm}
\item Develop motor skills learning algorithm by incorporating real-time estimates.
\end{itemize}

\end{block}

\end{column}

\begin{column}{0.47\textwidth}

\begin{block}{References}

\nocite{*}
\scriptsize{
\bibliographystyle{unsrt}

\begin{thebibliography}{10}
\bibitem{tamei}
T. Tamei, T. Matsubara, A. Rai, and T. Shibata, ``Reinforcement learning of clothing assistance with a dual-arm robot'' in \emph{Proc. of the 11th IEEE-RAS Int. Conf. on Humanoid Robots, pp. 733-738,} 2011.

\bibitem{topology}
E. S. L. Ho and T. Komura, ``Character motion synthesis by topology coordinates,'' in \emph{Proc. of EUROGRAPHICS2009, vol. 28, no. 2, pp. 299-308,} 2009.

\bibitem{ellipse}
M. C. Shakarji, ``Least-squares fitting algorithms of the NIST algorithm testing system.'', in \emph{Journal of Research-National Institute of Standards and Technology 103,} 1998.
\end{thebibliography}}

\end{block}

\end{column}

\begin{column}{.02\textwidth}\end{column}

\end{columns}

\end{frame}

\end{document}

### 2016-12-05 Poster/poster.tex
[minor] edit template file with current content

@@
 \begin{columns}[t]
-
   \begin{column}{.02\linewidth}\end{column}
@@
   \begin{column}{0.47\linewidth}
-
     \begin{block}{Introduction}
-
       \begin{center}
-\textbf{Objective}: Robotic Clothing Assistance is a challenging problem involving \emph{interaction with Human and Clothing material}.
-In this study, we propose a method for the realtime estimation of the Human-Cloth topological relationship using a depth sensor.
+        \textbf{Abstract}: Motor-skill learning for complex robotic tasks is a challenging problem due to the high task variability.
+Robotic clothing assistance is one such challenging problem that can greatly improve the quality-of-life for the elderly and disabled.
+In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent space learning.
+We implement our framework in a practical setting with a dual-arm robot performing clothing assistance tasks.
+We demonstrate that performing motor-skills learning in such task specific latent spaces outperforms learning in the high-dimensional joint configuration space of the robot.
+Furthermore, our framework can also be used as a tool for learning from demonstration to impart novel skills to the robot.
       \end{center}
-
-\begin{columns}[t]
-
-\begin{column}{0.55\linewidth}
-
-~\\
-Tamei et. al.[1] developed clothing assistance robot to perform T-shirt clothing task:\\~\\
-\begin{itemize}
-\item \textbf{Reinforcement Learning} scheme used to acquire motor skills for cloth handling.\\~\\
-\item \textbf{Low dimensional represetations} of state and policy used to ensure fast learning time.
-\end{itemize}
-
-\end{column}
-
-\begin{column}{0.45\linewidth}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\textwidth]{mannequin.png}
-\end{figure}
-
-\end{column}
-
-\end{columns}
-
     \end{block}
-
   \end{column}
@@
   \begin{column}{0.47\linewidth}
-
     \begin{block}{Problem Decription}
 
-\begin{columns}[t]
-
-\begin{column}{0.5\linewidth}
-\textbf{Topology Coordinates}[2] represents interaction between two curves using low dimensional variables:
-
-\begin{itemize}
-\item \textbf{Writhe}, $w \in \mathbb{R}$
-\vspace{3mm}
-\item \textbf{Center}, $\text{c} = [c_1,c_2] \in \mathbb{R}^2$
-\vspace{3mm}
-\item \textbf{Density}, $d \in \mathbb{R}$
-\end{itemize}
-
-\vspace{3mm}
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{humancloth.pdf}
-\end{figure}
-
-\end{column}
-
-\begin{column}{0.5\linewidth}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{topology.png}
-\end{figure}
-
-\vspace{3mm}
-
-\centering \textbf{Human-Cloth relationship} low dimensionally represented using topology coordinates:
-
-\begin{itemize}
-\item T-shirt collar and Mannequin's Head
-\vspace{2mm}
-\item T-shirt collar and Mannequin's Body
-\vspace{2mm}
-\item T-shirt sleeve and Mannequin's Shoulder
-\end{itemize}
-
-\end{column}
-
-\end{columns}
-
     \end{block}
-
   \end{column}
@@
   \begin{column}{.02\linewidth}\end{column}
-
 \end{columns}
@@
 \begin{columns}[t]
-
   \begin{column}{0.02\linewidth}\end{column}
@@
   \begin{column}{0.96\linewidth}
-
-\begin{alertblock}{Proposed Method}
-
+    \begin{alertblock}{Motor-skill Representations in Latent Spaces}
       \begin{columns}[t]
-
         \begin{column}{0.24\linewidth}
-
-\centering \textbf{Experimental Setup}
-
-\begin{itemize}
-\item Human-Cloth relationship observed using \textbf{Kinect depth sensor}.\\~\\
-\item \textbf{Color markers} used on T-shirt and Mannequin to ensure robust estimation.
-\end{itemize}
-
-\vspace{6mm}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{setup.png}
-\end{figure}
+          \centering \textbf{Learning from Demonstration}
 
@@
         \begin{column}{0.48\linewidth}
-
-\centering \textbf{Proposed Method}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth,height=0.22\textheight]{method.pdf}
-\end{figure}
+          \centering \textbf{Proposed Framework}
 
@@
         \begin{column}{0.24\linewidth}
-
-\centering \textbf{T-shirt state estimation}
-
-\vspace{3mm}
-
-\begin{itemize}
-\item T-shirt collar and sleeve shape compactly represented using \textbf{Ellipses}.\\~\\
-\item Generic 3D conic fitting approach proposed Shakarji et. al.~\cite{ellipse} used for this purpose.
-\end{itemize}
-
-\vspace{3mm}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{collarestimates.pdf}
-\end{figure}
+          \centering \textbf{Reinforcement Learning Formulation}
 
         \end{column}
+      \end{columns}
+    \end{alertblock}
+  \end{column}
 
+  \begin{column}{0.02\linewidth}\end{column}
 \end{columns}
-
-\end{alertblock}
 
@@
 
-\centering \textbf{Robustness of Topology Space to Measurement Noise}\\
+\centering \textbf{}\\
 \vspace{3mm}

### 2016-12-07 Poster/poster.tex
[major] prepare poster for NIPS

@@
 \usepackage[orientation=portrait,size=a0,scale=1.4]{beamerposter}
+
 \usetheme{I6pd2}
+\usepackage{xcolor}
+\usepackage{caption}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{booktabs}
+\usepackage{tcolorbox}
+\graphicspath{{Images/}}
 \usepackage[english]{babel}
+\usepackage[tight,footnotesize]{subfigure}
 \usepackage{amsmath,amsthm,amssymb,latexsym}
-\usepackage{booktabs}
-\usepackage{caption}
-\usepackage{tcolorbox}
-\usepackage{xcolor}
-\graphicspath{{Images/}}
-\usepackage[tight,footnotesize]{subfigure}
+
+\usepackage{tikz}
+\usetikzlibrary{arrows,shapes,backgrounds,shapes.misc,fit,bayesnet}
+
+\setbeamerfont{footnote}{size=\tiny}
 
@@
       \begin{center}
-        \textbf{Abstract}: Motor-skill learning for complex robotic tasks is a challenging problem due to the high task variability.
-Robotic clothing assistance is one such challenging problem that can greatly improve the quality-of-life for the elderly and disabled.
-In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent space learning.
-We implement our framework in a practical setting with a dual-arm robot performing clothing assistance tasks.
-We demonstrate that performing motor-skills learning in such task specific latent spaces outperforms learning in the high-dimensional joint configuration space of the robot.
-Furthermore, our framework can also be used as a tool for learning from demonstration to impart novel skills to the robot.
+        \textbf{Objective}: Motor-skill learning for complex robotic tasks is challenging due to high task variability.
+Robotic clothing assistance is one problem that can improve the quality-of-life for elderly and disabled.
+In this study, we propose a representation to encode task specific motor-skills using Bayesian nonparametric latent space learning and rely on this for data-efficient reinforcement learning.
       \end{center}
+
+      \begin{columns}[t]
+        \begin{column}{0.6\linewidth}
+          \vskip0.5em
+          Tamei et. al. \footnotemark[1] developed clothing assistance framework for dual-arm robots:\vskip0.5em
+          \begin{itemize}
+            \item \textbf{Reinforcement Learning} to acquire motor skills.
+            \item Policy representation inefficient for generalization.
+          \end{itemize}
+
+          \begin{figure}
+            \centering
+            \includegraphics[height=0.06\textheight, width=0.8\textwidth]{policy.png}
+          \end{figure}
+        \end{column}
+
+        \begin{column}{0.4\linewidth}
+          \begin{figure}
+            \centering
+            \includegraphics[width=0.9\textwidth]{setup.png}
+          \end{figure}
+        \end{column}
+      \end{columns}
     \end{block}
@@
   \begin{column}{0.47\linewidth}
-    \begin{block}{Problem Decription}
+    \begin{block}{Latent Space Reinforcement Learning}
+      \begin{columns}[t]
+        \begin{column}{0.65\textwidth}
+          Combine Reinforcement Learning (RL) with Dimensionality Reduction (DR):
+          \begin{itemize}
+            \item Tractable search space reduces learning time.
+            \item Latent space can model task space constraints.
+          \end{itemize}
+        \end{column}
 
+        \begin{column}{0.35\textwidth}
+          \vskip-2em
+          \begin{figure}
+            \centering
+            \includegraphics[width=0.8\textwidth]{latentrl.png}
+            \caption*{\footnotesize{Bitzer et al., 2010 \footnotemark[2]}}
+          \end{figure}
+        \end{column}
+      \end{columns}
+
+      \begin{columns}[t]
+        \begin{column}{0.3\textwidth}
+          \vskip-1em
+          \begin{figure}
+            \centering
+            \caption*{Bitzer et al., 2010 \footnotemark[2]}
+            \includegraphics[width=0.8\textwidth]{bitzer.png}
+          \end{figure}
+
+          \begin{figure}
+            \centering
+            \caption*{Luck et al., 2014 \footnotemark[3]}
+            \includegraphics[width=0.8\textwidth]{luck.png}
+          \end{figure}
+        \end{column}
+
+        \begin{column}{0.6\textwidth}
+          \vskip0.5em
+          Existing frameworks have limitations:\vskip0.5em
+          \begin{itemize}
+            \item \textbf{Bitzer et al., 2010 \footnotemark[2]}: DR as preprocessing before RL
+              \begin{itemize}
+                \item MAP estimate of latent space leads to overfitting.\\~\\
+              \end{itemize}
+              \item \textbf{Luck et al., 2014 \footnotemark[3]}: Combine DR and RL under EM-formulation
+                \begin{itemize}
+                  \item Linear model used for DR can be insufficient for complex tasks.
+                \end{itemize}
+          \end{itemize}
+        \end{column}
+      \end{columns}
     \end{block}
@@
   \begin{column}{0.96\linewidth}
-    \begin{alertblock}{Motor-skill Representations in Latent Spaces}
+    \begin{block}{Proposed Method}
       \begin{columns}[t]
-        \begin{column}{0.24\linewidth}
-          \centering \textbf{Learning from Demonstration}
+        \begin{column}{0.33\linewidth}
+          \centering \large \textbf{Bayesian Gaussian Process LVM}
 
+          \begin{columns}
+            \begin{column}{0.8\textwidth}
+              \begin{itemize}
+                \item Latent variable model (Titsias et al., 2010 \footnotemark[4]):
+                  \begin{equation*}
+                    \mathbf{y} = f(\mathbf{x}) + \epsilon,~~\epsilon \in \mathcal{N}(\mathbf{0}, \sigma^2 \mathbf{I})
+                  \end{equation*}
+                  \begin{itemize}
+                    \item $\mathbf{y} \in \mathbb{R}^D$: Observed Variable
+                    \item $\mathbf{x} \in \mathbb{R}^Q (Q \ll D)$: Unknown latent variable
+                    \item $f: \mathbf{x} \rightarrow \mathbf{y}$: Gaussian Process mapping
+                  \end{itemize}
+              \end{itemize}
+            \end{column}
+            \begin{column}{0.2\textwidth}
+              \begin{figure}
+                \centering
+                \includegraphics[width=\textwidth]{lvm.png}
+              \end{figure}
+            \end{column}
+          \end{columns}
+          \begin{figure}
+            \centering
+            \includegraphics[width=0.8\textwidth]{nonlinearMap.png}
+          \end{figure}
+
+          \begin{itemize}
+            \item \textbf{Bayesian Inference}: Posterior distribution on the latent space.
+              \begin{equation*}
+                p(\mathbf{Y}) = \int_{\mathbf{X}} p(\mathbf{Y}|\mathbf{X}) p(\mathbf{X}) d\mathbf{X}
+              \end{equation*}
+            \item Automatic dimensionality reduction possible using ARD kernel \footnotemark[5]:
+                \begin{equation*}
+                  k(x,x') = \sigma_f^2 \exp \left( - \frac{1}{2} \Sigma_{q=1}^Q \mathbf{w_q} (x_q - x_q')^2 \right)
+                \end{equation*}
+          \end{itemize}
         \end{column}
 
-        \begin{column}{0.48\linewidth}
-          \centering \textbf{Proposed Framework}
+        \begin{column}{0.34\linewidth}
+          \centering \large \textbf{Motor-skills in Latent Space}
+          \begin{tcolorbox}[width=0.9\textwidth, colback=red!75!white, colframe=red!5!white]
+            \small \centering \textcolor{white}{\textbf{Problem}: Design of robust motor-skills framework is crucial for real-world implementation on low-cost robots.}
+          \end{tcolorbox}
 
+          \begin{tcolorbox}[width=0.9\textwidth, colback=blue!75!white, colframe=blue!5!white]
+            \small \centering \textcolor{white}{\textbf{Solution}: Use Bayesian nonparametric dimensionality reduction for efficient learning of clothing skills.}
+          \end{tcolorbox}
+          \vskip0.5em
+          \begin{figure}
+            \centering
+            \includegraphics[width=\textwidth]{overview.png}
+          \end{figure}
+
+          \begin{itemize}
+            \item \small \textbf{Learning from Demonstration}: BGPLVM latent space as user-friendly controller for kinesthetic demonstration.
+            \item \small \textbf{Reinforcement Learning}: Apply reinforcement learning with BGPLVM latent space as action space.
+          \end{itemize}
         \end{column}
 
-        \begin{column}{0.24\linewidth}
-          \centering \textbf{Reinforcement Learning Formulation}
+        \begin{column}{0.33\linewidth}
+          \centering \large \textbf{Reinforcement Learning Formulation}
 
+          \begin{itemize}
+            \item Cross Entropy Method (CEM)\footnotemark[6] for policy improvement:
+              \begin{equation*}
+                \begin{array}{c}
+                  \theta^* \sim \mathcal{N} (\theta | \mathbf{\mu}^*, \Sigma^{*})\\
+                  \mathbf{\mu}^* := \text{mean}(\text{argmax}~ \theta_{\text{old}}),~~\Sigma^{*} := \text{var}(\text{argmax} ~\theta_{\text{old}})
+                \end{array}
+              \end{equation*}
+            \item \textbf{Policy representation}: Dynamic Movement Primitives (DMP) \footnotemark[7]:
+              \begin{equation*}
+                \begin{array}{c}
+                  \tau \ddot{x} = K(g-x) -D\dot{x} + (g - x_0)f\\
+                  \displaystyle f(s) = \frac{\Sigma_i w_i \psi_i(s)s}{\Sigma_i \psi_i(s)}, \text{ where } \tau \dot{s} = - \alpha s
+                \end{array}
+              \end{equation*}
+            \item \textbf{Reward function}: Distance from desired Via-points to current policy:
+              \begin{equation*}
+                R(\pi(\theta)) = \Sigma_{i = 1}^{n_{\text{dims}}} \Sigma_{j = 1}^{n_{\text{via}}} \| V_{i,j} - \pi_i(\theta, t_{i,j}) \|^2
+              \end{equation*}
+          \end{itemize}
+
+          \begin{figure}
+            \centering
+            \includegraphics[width=0.8\textwidth]{trajectory.png}
+          \end{figure}
         \end{column}
       \end{columns}
-    \end{alertblock}
+    \end{block}
   \end{column}
@@
 
-\begin{alertblock}{Experimental Results}
+\begin{block}{Experimental Results}
 
   \begin{columns}[t]
+    \begin{column}{0.02\linewidth}\end{column}
 
-\begin{column}{0.49\linewidth}
+    \begin{column}{0.4\linewidth}
+      \centering \textbf{\large Generalization in Latent Space}\\
 
-\centering \textbf{}\\
-\vspace{3mm}
-\centering Robustness evaluated by comparing rewards obtained in \emph{Cartesian} and \emph{Topological} space.
+      \begin{columns}
+        \begin{column}{0.5\textwidth}
+          \begin{figure}
+            \centering
+            \includegraphics[width=0.9\textwidth]{jaError.png}
+          \end{figure}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+          \centering \textbf{Evaluation}: Reconstruction error of latent space with RMS Error \footnotemark[1]. \\
+          \vskip1em
+          \textbf{Dataset}: Clothing trajectories for 4 postures: Shoulder Angle $\in \{65^o,70^o,75^o,80^o\}$.
+        \end{column}
+      \end{columns}
+      \vskip1em
+      \begin{columns}
+        \begin{column}{0.36\textwidth}
+          \centering \textbf{PCA}
+          \vskip-0.5em
+          \begin{figure}
+            \centering
+            \includegraphics[width=\textwidth]{jaPCA.png}
+          \end{figure}
+        \end{column}
+        \begin{column}{0.36\textwidth}
+          \centering \textbf{GPLVM}
+          \vskip-0.5em
+          \begin{figure}
+            \centering
+            \includegraphics[width=\textwidth]{jaGPLVM.png}
+          \end{figure}
+        \end{column}
+        \begin{column}{0.36\textwidth}
+          \centering \textbf{BGPLVM}
+          \vskip-0.5em
+          \begin{figure}
+            \centering
+            \includegraphics[width=\textwidth]{jaBGPLVM.png}
+          \end{figure}
+        \end{column}
+      \end{columns}
+    \end{column}
+
+    \begin{column}{0.05\linewidth}\end{column}
+
+    \begin{column}{0.4\linewidth}
+      \centering \large \textbf{Effectivity of Search Space}\\
+
+      \begin{center}
+        Apply Reinforcement Learning in different action spaces with same formulation:
+      \end{center}
 
       \begin{itemize}
-\item Reward function in Cartesian Space:
-\begin{equation}
-r(o_t) = - \sqrt{(o_t - o_{target})^T \Sigma_o^{-1} (o_t - o_{target})}
-\end{equation}
-\item Reward function in Topology Space:
-\begin{equation}
-r(s_t) = - \sqrt{(s_t - s_{target})^T \Sigma_s^{-1} (s_t - s_{target})}
-\end{equation}
+        \item Parameters: $50 \times n_{\text{dims}}$ basis functions in DMP.
+        \item Cross Entropy Method: 50 rollouts per iteration.
+        \item Policy Update: 5 best rollouts per iteration
       \end{itemize}
-
+      \vskip-2em
       \begin{columns}[t]
-
-\begin{column}{0.49\linewidth}
-
-\centering Simulation Experiment
+        \begin{column}{0.35\textwidth}
+          \vskip2em
+          \begin{itemize}
+            \item BGPLVM search space performs best as it uses non-linear mapping and avoids overfitting. \vskip1em
+            \item PCA performs better than GPLVM as MAP estimation of GPLVM overfits to training data.
+          \end{itemize}
+        \end{column}
+        \begin{column}{0.65\textwidth}
           \begin{figure}
             \centering
-\includegraphics[width=\textwidth]{toptest.pdf}
+            \includegraphics[width=\textwidth]{rlcurve.png}
           \end{figure}
-
+        \end{column}
+      \end{columns}
     \end{column}
 
-\begin{column}{0.49\linewidth}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{reward.png}
-\end{figure}
-
-\end{column}
-
+    \begin{column}{0.02\linewidth}\end{column}
   \end{columns}
-
-\end{column}
-
-\begin{column}{0.49\linewidth}
-
-\centering \textbf{Accuracy of Proposed Method}
-
-\begin{itemize}
-\item Accuracy of proposed method evaluated by comparing our estimates with the ground truth.
-\vspace{3mm}
-\item \textbf{Ground Truth}: Motion Capture System with 7 IR Cameras, 10 IR Markers used on T-shirt collar and 1 IR Marker on T-shirt sleeve.
-\end{itemize}
-
-\begin{columns}[t]
-
-\begin{column}{0.59\linewidth}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{realtimeresults.pdf}
-\end{figure}
-
-\end{column}
-
-\begin{column}{0.39\linewidth}
-
-\begin{table}
-\centering
-\begin{center}
-\small
-\begin{tabular}{ccc}
-\toprule
-Parameter & Pearson Corr & RMS Error\\[0.2cm]
-\midrule
-$s_1$ topology & 0.9364 & 0.5955\\
-$s_2$ topology & 0.8857 & 0.1247\\
-$s_3$ topology & 0.9889 & 0.3529\\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-\vspace{4mm}
-
-\begin{itemize}
-\item Proposed method provides reliable estimates in real-time (28-30 fps).
-\vspace{3mm}
-\item However, there are jitters due to inherent noise in depth sensor.
-\end{itemize}
-
-\end{column}
-
-\end{columns}
-
-\end{column}
-
-\end{columns}
-
-\end{alertblock}
-
-\end{column}
-
-\begin{column}{0.02\textwidth}\end{column}
-
-\end{columns}
+\end{block}
 
@@
   \begin{column}{0.47\textwidth}
-
     \begin{block}{Future Work}
-
       \begin{itemize}
-\item Bayesian estimation of Human-Cloth relationship using \emph{Proprioceptive} information of robot.
-\vspace{3mm}
-\item Develop motor skills learning algorithm by incorporating real-time estimates.
+        \item \textbf{Immediate Goal}: BGPLVM for learning clothing assistance skills:
+          \begin{itemize}
+            \item Use robot proprioceptive and depth sensor readings as state representation.
+            \item Incorporate dynamics prior and task-specific covariance functions.
           \end{itemize}
-
+          \vskip0.5em
+        \item \textbf{Long Term Goal}: Combine policy search RL and BGPLVM:
+          \begin{itemize}
+            \item EM-formulation with BGPLVM parameters as latent variables.
+            \item Modify variational inference framework for RL setting.
+          \end{itemize}
+        \end{itemize}
     \end{block}
-
   \end{column}
@@
   \begin{column}{0.47\textwidth}
-
     \begin{block}{References}
-
       \nocite{*}
@@
         \bibitem{tamei}
-T.
-Tamei, T.
-Matsubara, A.
-Rai, and T.
-Shibata, ``Reinforcement learning of clothing assistance with a dual-arm robot'' in \emph{Proc.
-of the 11th IEEE-RAS Int.
-Conf.
-on Humanoid Robots, pp.
-733-738,} 2011.
+        Tamei, T. et al., ``Reinforcement learning of clothing assistance'', in \emph{IEEE-RAS Humanoids 2011}
 
-\bibitem{topology}
-E. S. L. Ho and T. Komura, ``Character motion synthesis by topology coordinates,'' in \emph{Proc. of EUROGRAPHICS2009, vol. 28, no. 2, pp. 299-308,} 2009.
+        \bibitem{bitzer}
+        Bitzer, S. et al., ``Using dimensionality reduction in reinforcement learning'' in \emph{IEEE/RSJ IROS, 2010}
 
-\bibitem{ellipse}
-M. C. Shakarji, ``Least-squares fitting algorithms of the NIST algorithm testing system.'', in \emph{Journal of Research-National Institute of Standards and Technology 103,} 1998.
+        \bibitem{luck}
+        Luck, K. S. et al., ``Latent space policy search for robotics'' in \emph{IEEE/RSJ IROS, 2014}
+
+        \bibitem{bgplvm}
+        Titsias, M. K. et al., ``Bayesian Gaussian Process Latent Variable Model'', in \emph{AISTATS 2011}
+
+        \bibitem{gp}
+        Williams, C. et al., ``Gaussian processes for machine learning'' by \emph{MIT Press 2006}.
+
+        \bibitem{cem}
+        De Boer, P. et al., ``A tutorial on the cross-entropy method'' in \emph{Annals of operations research 2005}.
+
+        \bibitem{dmp}
+        Ijspeert, A. J. et al., ``Dynamical movement primitives: learning attractor models for motor behaviors'', in \emph{Neural computation 2013}.
       \end{thebibliography}}
-
     \end{block}
-
   \end{column}
@@
   \begin{column}{.02\textwidth}\end{column}
-
 \end{columns}

### 2017-01-10 Highlights/slides.tex
[major] layout for nips presentation

% Presentation for RSJ Annual Conference
% Author: Nishanth Koganti
% Date: 2017/01/10

\documentclass[11pt,
               hyperref={colorlinks,citecolor=pink,linkcolor=red,urlcolor=blue}
               ]{beamer}

\usetheme{default}

\usefonttheme{default}
\usecolortheme{whale}

\usepackage{bm}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{subfig}
\usepackage{caption}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{lmodern}
\usepackage{hyperref}
\usepackage{fancybox}
\usepackage{amsfonts}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{subcaption}
\graphicspath{{./Images/}}
\usepackage[absolute,overlay]{textpos}
\usetikzlibrary{arrows,shapes,backgrounds,shapes.misc,fit}

\everymath{\displaystyle}
\setbeamerfont{footnote}{size=\tiny}

\setbeamercovered{invisible}
\setbeamercolor{footline}{fg=black}
\setbeamertemplate{blocks}[rounded]
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]
\setbeamerfont{footline}{size=\footnotesize}
\setbeamerfont{block body}{size=\footnotesize}
\setbeamertemplate{frametitle}[default][center]
\setbeamerfont{block title}{size=\footnotesize}
\setbeamercolor*{block body}{fg= white,bg= blue!60}
\setbeamercolor*{block title}{fg= white,bg= blue!70}
\setbeamerfont{block body example}{size=\footnotesize}
\setbeamerfont{block title example}{size=\footnotesize}
\setbeamercolor*{block body example}{fg=white,bg= red!60}
\setbeamercolor*{block title example}{fg=white,bg= red!70}

\newcommand{\highlight}[2][yellow]{\colorbox{#1}{$\displaystyle #2$}}
\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}

\tikzset{square arrow/.style={to path={-- ++(0,-.25)  -| (\tikztotarget) \tikztonodes},below,pos=.25}}

\tikzstyle{line} = [draw, ultra thick, -latex']
\tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text centered, rounded corners, minimum height=3em, font=\scriptsize]

\tikzstyle{block0} = [rectangle, draw, fill=blue!20, text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{block1} = [rectangle, draw, fill=green!20, text width=5em, text centered, rounded corners, minimum height=4em,minimum width=12em]
\tikzstyle{block2} = [rectangle, draw, text width=5em, text centered, rounded corners, minimum height=6em,minimum width=35em]
\tikzstyle{line0} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm, minimum height=2em]

\newenvironment{reference}[2]{\begin{textblock*}{\textwidth}(#1,#2) \footnotesize\it\bgroup\color{red!50!black}}{\egroup\end{textblock*}}

\title[NIPS Highlights]{Highlights of NIPS 2016}
\author[Nishanth]{Nishanth Koganti}
\institute[NAIST]{Research Student, Shibata Lab \\ \tiny{Kyushu Institute of Technology}}
\date{January 13, 2017}

\titlegraphic{\includegraphics[height=2cm]{nipsLogo.png}\hspace*{1.5cm}\includegraphics[height=2cm]{nipsWords.png}\hspace*{1.5cm}\includegraphics[height=2cm]{barcelonaLogo.jpg}}

\begin{document}

  \begin{frame}[noframenumbering]
    \titlepage
  \end{frame}

  \section{NIPS Conference}

  \subsection{History}

  \begin{frame}
    \frametitle{Neural Information Processing Systems}

    \begin{itemize}
      \item Largest International Conference on Machine Learning.
      \item Held since 1987 and initially focused on Neural Network research.
    \end{itemize}

    \begin{figure}
      \centering
      \includegraphics[width=0.65\textwidth]{terryLaw.jpg}
      \caption*{Terry's Law on NIPS Participants \footnotemark[1]}
    \end{figure}

    \footnotetext[1]{Proposed by \href{http://www.salk.edu/scientist/terrence-sejnowski/}{Terrence Sejnowski}, NIPS 2016 Chair}
  \end{frame}

  \subsection{Statistics}

  \begin{frame}
    \frametitle{NIPS 2016: Topics Convered}

    \begin{itemize}
      \item 568 Papers (46 oral) accepted among 2400 papers with 6 reviewers per submission.
      \item Most popular topics were Deep Learning (1 in 4) with application to Computer Vision (1 in 10).
    \end{itemize}

    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{nipsTopics1.png}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{nipsTopics2.png}
      \caption*{Long-tailed Distribution of Topics \footnotemark[1]}
    \end{figure}

    \footnotetext[1]{\href{http://www.tml.cs.uni-tuebingen.de/team/luxburg/misc/nips2016/index.php}{NIPS 2016 Review Process}}
  \end{frame}

  \subsection{Impressions}

  \begin{frame}
    \frametitle{NIPS 2016: Impressions}

    \begin{itemize}
      \item
    \end{itemize}

    %\footnotetext[1]{}
  \end{frame}

  \section{Deep Learning}

  \subsection{Generative Adversarial Networks}

  \begin{frame}
    \frametitle{Generative Adversarial Networks: Overview}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{GAN: Tutorials}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{GAN: Applications}

    %\footnotetext[1]{}
  \end{frame}

  \subsection{Bayesian Deep Learning}

  \begin{frame}
    \frametitle{Bayesian Deep Learning}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{Deep Learning and Graphical Models}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{Dropout as a Bayesian Approximation}

    %\footnotetext[1]{}
  \end{frame}

  \subsection{Deep Reinforcement Learning}

  \begin{frame}
    \frametitle{Deep Reinforcement Learning: Tutorials}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{Deep Reinforcement Learning: Workshop}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{Value Iteration Networks}

    %\footnotetext[1]{}
  \end{frame}

  \subsection{Recurrent Neural Networks}

  \begin{frame}
    \frametitle{Recurrent Neural Networks: Overview}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{RNN: Symposia}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{RNN: Applications}

    %\footnotetext[1]{}
  \end{frame}

  \subsection{Meta-learning}

  \begin{frame}
    \frametitle{Learning to learn: Meta-learning}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{One-shot Learning}

    %\footnotetext[1]{}
  \end{frame}

  \subsection{Neural Turing Computers}

  \begin{frame}
    \frametitle{Neural Turing Machines: Overview}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{NTM: Applications}

    %\footnotetext[1]{}
  \end{frame}

  \section{General Topics}

  \begin{frame}
    \frametitle{Nuts and Bolts of ML}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{General AI: Unsupervised Learning}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{Boston Robotics Demo}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{Efficient seeding for K-Means}

    %\footnotetext[1]{}
  \end{frame}

  \begin{frame}
    \frametitle{RocketAI: The AI Bubble}

    %\footnotetext[1]{}
  \end{frame}

  \section{Visitng Barcelona}

  \begin{frame}
    \frametitle{Sightseeing in Barcelona}

  \end{frame}

  \begin{frame}[noframenumbering]

    \begin{center}
      \Huge{Thank you!}\\~\\
      \huge{Questions or Comments?}
    \end{center}
  \end{frame}

\end{document}

### 2017-01-11 Highlights/slides.tex
[major] updates to highlights presentation

@@
       \item Largest International Conference on Machine Learning.
-      \item Held since 1987 and initially focused on Neural Network research.
+      \item Held since 1987 and focuses on AI, ML, Mathematics, Statistics.
     \end{itemize}
@@
     \begin{itemize}
-      \item
+      \item \textbf{Strong industry presence}: Google Deepmind, Facebook AI Research, OpenAI etc
+      \item \textbf{Philosophical panel discussions}: Healthy debates on general trends and future of community.
+      \item \textbf{Emphasis on Deep Learning}: Studies on applications as well as \textbf{theoretical foundations} of Deep Learning.
     \end{itemize}
+    \vskip-1em
+    \begin{columns}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{Tutorial on GANs\footnotemark[1]}
+          \includegraphics[width=\textwidth]{goodfellowTutorial.jpg}
+        \end{figure}
+      \end{column}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{Conference Venue\footnotemark[1]}
+          \includegraphics[width=\textwidth]{conferenceVenue.jpeg}
+        \end{figure}
+      \end{column}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{Boston Dynamics\footnotemark[1]}
+          \includegraphics[width=\textwidth]{bostonDyn.png}
+        \end{figure}
+      \end{column}
+    \end{columns}
 
-    %\footnotetext[1]{}
+    \footnotetext[1]{\href{http://blog.aylien.com/wp-content/uploads/2016/12/full_conference_hall_gan_tutorial.jpg}{Ian Goodfellow's Tutorial on GAN}, \href{https://cdn-images-1.medium.com/max/800/1*l6ILYWwpF9A1KkqK6hwuaA.jpeg}{Image of Conference Venue}, \href{http://sebastianruder.com/content/images/2016/12/boston_dynamics_spot.png}{Boston Dynamics Spot Demo}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \begin{equation*}
+      \text{min}_G~\text{max}_D~V(D,G) = \mathbb{E}_{x \sim p_{\text{data}}(x)} \log D(x) + \mathbb{E}_{x \sim p_{x}(z)} \log (1 - D(G(z)))
+    \end{equation*}
+    \vskip-1em
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.75\textwidth]{gan1.png}
+    \end{figure}
+
+    \footnotetext[1]{Goodfellow et al. ``Generative adversarial nets'' in \emph{NIPS 2016}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item \href{http://www.iangoodfellow.com/slides/2016-12-04-NIPS.pdf}{Tutorial on GANs by Ian Goodfellow}:
+        \begin{itemize}
+          \item Applications to generate realistic images.
+          \item Learn latent manifolds encoding task variability.
+          \item Theoretical challenges for training models.
+        \end{itemize}
+      \item \href{https://github.com/soumith/ganhacks}{Tutorial on Practical Tips by Soumith Chintala}:
+        \begin{itemize}
+          \item Tips on data preprocessing, network design and optimizers.
+        \end{itemize}
+    \end{itemize}
+    \vskip-1em
+    \begin{columns}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{Manifold on MNIST\footnotemark[1]}
+          \includegraphics[width=0.9\textwidth]{gan2.png}
+        \end{figure}
+      \end{column}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{Tutorial by Chintala\footnotemark[1]}
+          \includegraphics[width=0.9\textwidth]{chintalaTutorial.png}
+        \end{figure}
+      \end{column}
+    \end{columns}
+
+    \footnotetext[1]{\href{http://www.iro.umontreal.ca/~bengioy/talks/dlss-12aug2015.pdf}{Presentation at DLSS 2015}, \href{https://github.com/soumith/ganhacks}{How to Train Your GAN}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \begin{columns}
+      \begin{column}{0.5\textwidth}
+        \centering
+        Learning What and Where\footnotemark[1]
+
+        \begin{figure}
+          \centering
+          \includegraphics[width=0.75\textwidth, height=0.25\textheight]{ganApp1.png}
+        \end{figure}
+        \vskip-1em
+        \begin{itemize}
+          \item Use text input to generate novel images.
+          \item Architecture for location and content controllable synthesis.
+        \end{itemize}
+      \end{column}
+
+      \begin{column}{0.5\textwidth}
+        \centering
+        Information Maximizing GANs\footnotemark[2]
+
+        \begin{figure}
+          \centering
+          \includegraphics[width=0.9\textwidth]{ganApp2.png}
+        \end{figure}
+        \vskip-1em
+        \begin{itemize}
+          \item Unsupervised learning of \textbf{disentagled} representations.
+          \item Maximizes mutual information between latent variables and observations.
+        \end{itemize}
+      \end{column}
+    \end{columns}
+
+    \footnotetext[1]{Reed et al., ``Learning what and where to draw'' in \emph{NIPS 2016}}
+    \footnotetext[2]{Chen et al., ``Infogan: Interpretable representation learning by information maximizing generative adversarial nets'' in \emph{NIPS 2016}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item \href{http://bayesiandeeplearning.org/slides/nips16bayesdeep.pdf}{History of Bayesian Neural Networks by Zoubin Gharamani\footnotemark[1]}:
+        \begin{itemize}
+          \item Derives from Gaussian Process Regression
+          \item Feature selection possible using Automatic Relevance Determination
+          \item Current research on variational inference techniques
+        \end{itemize}
+    \end{itemize}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=\textwidth]{bnn1.png}
+    \end{figure}
+
+    \footnotetext[1]{Keynote Talk by Zoubin Gharamani}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \centering Combined training of Graphical Models and Deep Neural Networks using Variational Inference\footnotemark[1]
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.8\textwidth]{bnn2.png}
+    \end{figure}
+
+    \footnotetext[1]{Johnson et al., ``Composing graphical models with neural networks for structured representations and fast inference'' in \emph{NIPS 2016}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \centering Dropout is equivalent to variational inference in Deep GPs with specific kernel function
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=\textwidth]{bnn4.png}
+    \end{figure}
+
+    \footnotetext[1]{Gal et al., ``A theoretically grounded application of dropout in recurrent neural networks'' in \emph{NIPS 2016}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item \href{http://people.eecs.berkeley.edu/~pabbeel/nips-tutorial-policy-optimization-Schulman-Abbeel.pdf}{Deep RL through Policy Optimization\footnotemark[1]}:
+        \begin{itemize}
+          \item History of policy search RL
+          \item Recent trends with DNN policies presented
+          \item Future work on \textbf{Skill Transfer} and \textbf{Exploration}
+        \end{itemize}
+      \item \href{http://rll.berkeley.edu/deeprlcourse/docs/nuts-and-bolts.pdf}{Nuts and Bolts of Deep RL Research\footnotemark[2]}:
+        \begin{itemize}
+          \item Tips on implementing existing algorithms
+          \item Problem formulation for novel tasks
+        \end{itemize}
+    \end{itemize}
+
+    \begin{columns}
+      \begin{column}
+        \begin{figure}
+          \centering
+          \includegraphics[width=0.8\textwidth]{}
+        \end{figure}
+      \end{column}
+      \begin{column}
+        \begin{figure}
+
+        \end{figure}
+      \end{column}
+    \end{columns}
+    \footnotetext[1]{Tutorial by Pieter Abbeel and John Schulman}
+    \footnotetext[2]{Tutorial by John Schulman in Deep RL Workshop}
   \end{frame}

### 2017-01-12 Highlights/slides.tex
[major] second update to slides for highlights presentation

@@
       \centering
-      \includegraphics[width=0.8\textwidth]{bnn2.png}
+      \includegraphics[width=0.75\textwidth]{bnn2.png}
     \end{figure}
@@
         \begin{itemize}
-          \item History of policy search RL
-          \item Recent trends with DNN policies presented
+          \item History of policy search RL focusing on Deep RL
           \item Future work on \textbf{Skill Transfer} and \textbf{Exploration}
@@
 
-    \begin{columns}
-      \begin{column}
     \begin{figure}
       \centering
-          \includegraphics[width=0.8\textwidth]{}
+      \includegraphics[width=0.8\textwidth]{drl1.png}
     \end{figure}
-      \end{column}
-      \begin{column}
-        \begin{figure}
 
-        \end{figure}
-      \end{column}
-    \end{columns}
     \footnotetext[1]{Tutorial by Pieter Abbeel and John Schulman}
@@
   \begin{frame}
-    \frametitle{Deep Reinforcement Learning: Workshop}
+    \frametitle{Opensource Environments}
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item OpenAI Universe\footnotemark[1]: Develop and implement RL algorithms on many environments.
+      \item Google Deepmind\footnotemark[2]: Simulator of 3D environments such as \textbf{Quake} and puzzle tasks.
+      \item Facebook TorchCraft\footnotemark[3]: Bridge between Torch Library and Starcraft Game.
+    \end{itemize}
+    \vskip-1.5em
+    \begin{columns}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{OpenAI Universe}
+          \includegraphics[width=0.8\textwidth]{drlApp1.png}
+        \end{figure}
+      \end{column}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{Google Deepmind}
+          \includegraphics[width=0.8\textwidth, height=0.25\textheight]{drlApp2.png}
+        \end{figure}
+      \end{column}
+      \begin{column}{0.33\textwidth}
+        \begin{figure}
+          \centering
+          \caption*{Facebook Starcraft}
+          \includegraphics[width=0.8\textwidth]{drlApp3.png}
+        \end{figure}
+      \end{column}
+    \end{columns}
+
+    \footnotetext[1]{\href{https://openai.com/blog/universe/}{OpenAI Universe Press Release}}
+    \footnotetext[2]{\href{https://deepmind.com/blog/open-sourcing-deepmind-lab/}{Google Deepmind Blog}}
+    \footnotetext[3]{\href{https://github.com/TorchCraft/TorchCraft}{TorchCraft Repository on Github}}
   \end{frame}
@@
   \begin{frame}
-    \frametitle{Value Iteration Networks}
+    \frametitle{Value Iteration Networks: Best Paper Award}
 
-    %\footnotetext[1]{}
+    \centering
+    Represent Bellman equations with Convolutional Neural Networks to learn Value function
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.8\textwidth]{drl2.png}
+    \end{figure}
+
+    \footnotetext[1]{Tamar et al., ``Value Iteration Networks'' in \emph{NIPS 2016}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item Papers on theoretical foundations (\textbf{better architectures}\footnotemark[1]) and applications (\textbf{Reinforcement Learning}\footnotemark[2]).
+      \item Strong correspondence to Neuroscience research (Fast and Slow time scales\footnotemark[3]).
+    \end{itemize}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.8\textwidth]{rnn1.png}
+    \end{figure}
+
+    \footnotetext[1]{}
   \end{frame}

### 2017-01-15 Highlights/slides.tex
[major] final final version of slides for highlights presentation

@@
     \footnotetext[1]{\href{http://blog.aylien.com/wp-content/uploads/2016/12/full_conference_hall_gan_tutorial.jpg}{Ian Goodfellow's Tutorial on GAN}, \href{https://cdn-images-1.medium.com/max/800/1*l6ILYWwpF9A1KkqK6hwuaA.jpeg}{Image of Conference Venue}, \href{http://sebastianruder.com/content/images/2016/12/boston_dynamics_spot.png}{Boston Dynamics Spot Demo}}
+  \end{frame}
+
+  \section{General Topics}
+
+  \begin{frame}
+    \frametitle{Nuts and Bolts of ML: Andrew Ng}
+
+    \begin{itemize}
+      \item Practical advice on machine learning research
+      \item Importance of human-level accuracy as metric
+      \item Techniques for collecting data to train models.
+    \end{itemize}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.8\textwidth]{ngKeynote.jpg}
+    \end{figure}
+
+    \footnotetext[1]{https://www.youtube.com/watch?v=F1ka6a13S9I}
+  \end{frame}
+
+  \begin{frame}
+    \frametitle{Predictive Learning: Yann Lecun}
+
+    \begin{figure}
+      \centering
+      \caption*{General AI means agents should have \textbf{Common Sense}\footnotemark[1]}
+      \includegraphics[width=0.8\textwidth]{generalAI2.png}
+    \end{figure}
+
+    \footnotetext[1]{\url{https://nips.cc/Conferences/2016/Schedule?showEvent=6197}}
+  \end{frame}
+
+  \begin{frame}
+    \frametitle{Predictive Learning: Yann Lecun}
+
+    \begin{figure}
+      \centering
+      \caption*{Key to General AI will be \textbf{Unsupervised Learning}\footnotemark[1]}
+      \includegraphics[width=0.8\textwidth]{generalAI1.png}
+    \end{figure}
+
+    \footnotetext[1]{\url{https://nips.cc/Conferences/2016/Schedule?showEvent=6197}}
+  \end{frame}
+
+  \begin{frame}
+    \frametitle{Intelligent Biosphere: Drew Purves}
+
+    \begin{figure}
+      \centering
+      \caption*{Intelligent life strongly depends on the environment and vice versa\footnotemark[1]}
+      \includegraphics[width=0.8\textwidth]{generalAI4.png}
+    \end{figure}
+
+    \footnotetext[1]{\url{https://nips.cc/Conferences/2016/Schedule?showEvent=6193}}
+  \end{frame}
+
+  \begin{frame}
+    \frametitle{Learning from the Brain: Saket Navlakha}
+
+    \begin{figure}
+      \centering
+      \caption*{Pruning of neural connections is crucial for Learning in Brain\footnotemark[1]}
+      \includegraphics[width=0.8\textwidth]{generalAI3.png}
+    \end{figure}
+
+    \footnotetext[1]{https://nips.cc/Conferences/2016/Schedule?showEvent=6192}
   \end{frame}
@@
       \centering
-      \includegraphics[width=0.8\textwidth]{rnn1.png}
+      \includegraphics[width=0.7\textwidth,height=0.45\textheight]{rnn2.png}
     \end{figure}
 
-    \footnotetext[1]{}
-  \end{frame}
-
-  \begin{frame}
-    \frametitle{RNN: Symposia}
-
-    %\footnotetext[1]{}
-  \end{frame}
-
-  \begin{frame}
-    \frametitle{RNN: Applications}
-
-    %\footnotetext[1]{}
+    \footnotetext[1]{http://papers.nips.cc/paper/6114-weight-normalization-a-simple-reparameterization-to-accelerate-training-of-deep-neural-networks}
+    \footnotetext[2]{https://arxiv.org/pdf/1611.02779.pdf}
+    \footnotetext[3]{http://papers.nips.cc/paper/6310-phased-lstm-accelerating-recurrent-network-training-for-long-or-event-based-sequences}
   \end{frame}
@@
   \begin{frame}
-    \frametitle{Learning to learn: Meta-learning}
+    \frametitle{Learning to (learn by gradient descent) by gradient descent}
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item Current optimizers are parametric and domain specific.
+      \item Attempt to learn the optimizer using RNN.
+    \end{itemize}
+    \begin{equation}
+      \theta_{t+1} := \theta_t + g(\Delta f(\theta_t), \phi)
+    \end{equation}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.9\textwidth]{metalearning1.png}
+    \end{figure}
+
+    \footnotetext[1]{Andrychowicz et al., ``Learning to learn by gradient descent by gradient descent'' in \emph{NIPS 2016}}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item Novel architectures proposed for one-shot learning.
+      \item The idea is a differentiable nearest neighbor classifier.
+    \end{itemize}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.8\textwidth]{oneshot1.png}
+    \end{figure}
+
+    \footnotetext[1]{Vinyals et al., ``Matching networks for one shot learning'' in \emph{NIPS 2016}}
   \end{frame}
 
-  \subsection{Neural Turing Computers}
+  \subsection{Differentiable Neural Computers}
 
   \begin{frame}
-    \frametitle{Neural Turing Machines: Overview}
+    \frametitle{Differentiable Neural Computers}
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item Neural Network with memory access (Read, Write, Link).
+      \item Used for skill transfer, reinforcement learning, reasoning etc.
+    \end{itemize}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.8\textwidth]{dnc.png}
+    \end{figure}
+
+    \footnotetext[1]{https://deepmind.com/blog/differentiable-neural-computers/}
   \end{frame}
 
-  \begin{frame}
-    \frametitle{NTM: Applications}
-
-    %\footnotetext[1]{}
-  \end{frame}
-
-  \section{General Topics}
-
-  \begin{frame}
-    \frametitle{Nuts and Bolts of ML}
-
-    %\footnotetext[1]{}
-  \end{frame}
-
-  \begin{frame}
-    \frametitle{General AI: Unsupervised Learning}
-
-    %\footnotetext[1]{}
-  \end{frame}
+  \section{Other Topics}
 
@@
 
-    %\footnotetext[1]{}
+    \begin{center}
+      \href{run:./Videos/BostonDyn.avi?loop}{\includegraphics[width=0.9\textwidth]{./Videos/BostonDyn.png}}
+    \end{center}
+
+    \footnotetext[1]{https://nips.cc/Conferences/2016/Schedule?showEvent=6194}
   \end{frame}
@@
   \begin{frame}
-    \frametitle{Efficient seeding for K-Means}
+    \frametitle{Classic Algorithms: KMeans and Matrix Completion}
 
-    %\footnotetext[1]{}
+    \begin{itemize}
+      \item Matrix completion: Non-convex objective for matrix completion has no spurious local minima \footnotemark[1]
+      \item Clustering: Seedings for large-scale k-means++ clustering \footnotemark[2]
+    \end{itemize}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.9\textwidth]{kmeans.png}
+    \end{figure}
+
+    \footnotetext[1]{http://papers.nips.cc/paper/6048-matrix-completion-has-no-spurious-local-minimum}
+    \footnotetext[2]{http://papers.nips.cc/paper/6478-fast-and-provably-good-seedings-for-k-means}
   \end{frame}
@@
 
-    %\footnotetext[1]{}
+    \centering Prank by NIPS organizers to warn about the AI Bubble
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.5\textwidth]{rocketai2.png}
+    \end{figure}
+
+    \footnotetext[1]{\url{https://medium.com/the-mission/rocket-ai-2016s-most-notorious-ai-launch-and-the-problem-with-ai-hype-d7908013f8c9}}
   \end{frame}
-
-  \section{Visitng Barcelona}
 
@@
 
+    \begin{columns}
+      \begin{column}{0.5\textwidth}
+        \begin{figure}
+          \centering
+          \includegraphics[width=\textwidth]{barc3.jpeg}
+        \end{figure}
+      \end{column}
+      \begin{column}{0.25\textwidth}
+        \begin{figure}
+          \centering
+          \includegraphics[width=\textwidth]{barc1.jpeg}
+        \end{figure}
+      \end{column}
+      \begin{column}{0.25\textwidth}
+        \begin{figure}
+          \centering
+          \includegraphics[width=\textwidth]{barc2.jpeg}
+        \end{figure}
+      \end{column}
+    \end{columns}
   \end{frame}
@@
       \Huge{Thank you!}\\~\\
-      \huge{Questions or Comments?}
+      \huge{Take home Messages}
+      \normalsize
+      \begin{itemize}
+        \item Deep Learning is no longer just a tool!
+        \item Emphasis is changing to General AI and Life long learning
+        \item Recurrent Neural Networks are very important!
+      \end{itemize}
     \end{center}
